Glycolipid MPIase is essential for the TAT (Twin-Arginine Translocation) pathway

Version 1

Abstract

TAT (Twin-Arginine Translocation) is a preprotein translocation system dedicated to membrane translocation of prefolded proteins in plants and bacteria. The TAT translocon, consisting of the TatABC subunits, drives translocation using proton motive force. However, there have been no reports on the successful reconstitution of the TAT system. In this report, we show that MPIase, a glycolipid that has known to catalyze membrane protein integration, is essential for the TAT system. Our findings in recombinant Escherichia coli demonstrate that overproducing TatABC increases MPIase levels and that depleting MPIase results in TAT precursor accumulation in the cytosol. Furthermore, co-reconstitution of MPIase with TatABC revealed the translocation activities of TAT substrates in a proton motive force-dependent manner. This is the first successful reconstitution of the TAT system and will be advantageous for understanding its mechanisms.

Introduction

The TAT (Twin-Arginine Translocation) system, found in plant chloroplasts, bacteria, and mitochondria of some protozoa and plants, is dedicated to the membrane translocation of presecretory proteins with a TAT signal sequence 1–4. This signal sequence contains the consensus motif of “SRRxFLK”, which includes the consecutive twin arginine residues that are essential for TAT translocation 5. The central h region of the TAT signal sequence is in average less hydrophobic than that of the Sec signal sequence 6. The c region of the TAT signal sequence often contains basic amino acids that have been shown to function as a Sec-avoidance motif 7, preventing recognition of the TAT substrates by the Sec pathway 8. Unlike the presecretory proteins that are translocated through the SecYEG or Sec61 translocon, TAT substrates maintain their folded structures during translocation. In many cases, TAT substrates fold and are transported with specific cofactors 5,9. TAT substrates are translocated through the translocation machinery formed by TatABC using the proton motive force (PMF) as the driving force 2,10–12. The TAT machinery consists of the TatA and TatC families, all of which are membrane proteins 13. In Escherichia coli, the TatA family includes TatA and TatB, as well as TatE 14–16. TatE is a functional homologue of TatA that is not essential for TAT translocation 14,17. TatA and TatB are structurally similar but functionally distinct, since both are essential for the TAT translocation 18–20. On the other hand, TatA and TatE have overlapping functions in the TAT system 14. All three proteins are integral membrane proteins containing a single transmembrane helix (TMH), followed by a cytoplasmic amphipathic helix (APH) 21,22. TatC is a polytopic membrane protein with six TMHs 23,24. TatC forms a complex with TatB at an equal molar ratio, which recognizes the signal peptides of TAT substrates 25–30. In contrast, TatA forms a homomultimer that associates with the TatBC complex with different stoichiometries depending on the size of the substrate 22,31–36. It is also reported that that a signal peptide alone can induce TatA multimerization 31. Binding of substrates to the TatBC complex causes TatA to assemble around the complex in a PMF-dependent manner 28,37. The TatA cluster then forms a translocon through which the TAT substrates can pass when bound to the TatBC complex 28,37–40. According to MD simulations, the short TMHs and APHs of the TatA cluster have been shown to thin the membranes, thereby destabilizing the lipid bilayer of the cell membranes and allowing the translocation of prefolded proteins 41,42. Once translocation is complete, the substrate is released from the TatABC complex, and the signal sequence is cleaved by membrane-located leader peptidase 10,43. Assembly of the TAT machinery is a transient process that occurs exclusively during substrate translocation 37,44. As mentioned above, extensive studies performed by numerous researchers have significantly advanced our understanding of the molecular mechanisms underlying of the TAT system, particularly following the identification of a Tat component in 1997 45. However, a reconstitution system for the TAT pathway has yet to be reported. Therefore, we hypothesized that a missing piece is necessary for successfully reconstituting the TAT system.

MPIase (Membrane Protein Integrase) is a glycolipid that catalyzes the integration of membrane proteins 46. Its glycan chain consists of approximately ten repeats of three N-acetylated amino sugars and is linked to diacylglycerol (DAG) by a pyrophosphate linker 46 (Fig. S1). This glycan chain has a molecular chaperon-like function, i.e., the glycan interacts directly with the substrate membrane proteins prior to insertion to prevent aggregation 46–48. The glycan is flexible on the membrane surface 49. DAG in the membrane bilayer blocks the disordered, spontaneous insertion by reducing the mobility of the acyl chain within the bilayers, while MPIase restores the mobility, allowing insertion with the substrate binding 49. In combination with the pyrophosphate linker, MPIase perturbs the membrane structure, enabling membrane proteins to insert into membranes. The negatively charged pyrophosphate of MPIase also attracts the positive charges near the TMH of the substrates 47,48.

MPIase can drive the membrane insertion of simple membrane proteins with a single TMH. MPIase also acts cooperatively with proteinaceous factors responsible for membrane insertion, such as the SecYEG translocase and the YidC insertase, in both Sec-dependent and - independent membrane insertion 50,51. MPIase is co-purified with SecYEG 52 and YidC 53., indicating that MPIase interacts with these proteins directly. MPIase also stimulates protein translocation by affecting the dimer structure of SecYEG 52,54,55. Thus, MPIase is generally involved in both membrane integration and translocation 56,57, leading us to hypothesize that MPIase is also involved in the TAT pathway.

Results

MPIase is involved in the TAT pathway in vivo

Recently, we found that CdsA and its paralogue YnbB, CDP-DAG synthases, are involved in MPIase biosynthesis and constructed the strain KS23 (ΔcdsA::cat ΔynbB) in which both the cdsA and ynbB genes were disrupted 58. The expression level of MPIase in this strain can be controlled by the addition of arabinose, since the cdsA gene is placed under the control of the araBAD promoter on a low-copy plasmid (pAra-CdsA). When arabinose is removed from the culture, MPIase becomes depleted (‘ΔMPIase’ conditions) 58. We investigated the effect of MPIase depletion on the TAT pathway by using KS23/pAra-CdsA and its parent strain (EK413) (Fig. 1). When the TAT substrate SufI was expressed in EK413, the signal sequence-cleaved mature form of the protein was observed, in addition to the precursor. The level of the mature form increased when TatABC was provided from the tet promoter on a plasmid (pTet-TatABC) in addition to chromosomal expression. This finding is consistent with a previous report 59. The amount of mature SufI did not change upon SufI induction, indicating that the level was saturated. A large part of untranslocated precursors was aggregated (Fig. S2). In marked contrast, no mature form was observed upon MPIase depletion in KS23/pAra-CdsA, even when TatABC was expressed from pTet-TatABC (Fig. 1A, top). Upon SufI induction, a smaller band marked by a star appeared in addition to the precursor; however, this band is likely a degradation product of the precursor that had not been translocated, as it is larger than the mature band. The addition of arabinose restored MPIase expression (Fig. 1A, middle left) and generated the mature form of SufI (Fig. 1A, top). These results strongly suggest that MPIase is involved in TAT translocation. Similar results were obtained, when using TorA, another TAT substrate: no mature form was observed upon MPIase depletion (Fig. 1A, bottom left). The low level of TAT substrates under the MPIase depletion is due to the fact that MPIase-depleted cells hardly proliferate. Although TorA translocation appears to be less efficient than SufI translocation, this might be because TorA’s folding efficiency is lower in the absence of trimethylamine N-oxide (TMAO), as previously reported 60. TatABC, which had been expressed prior to arabinose removal, was still present after MPIase depletion (see Fig. 2C). To confirm that the processing to the mature form represents the TAT translocation, we examined the TAT signal mutants SufI (KK) (Fig. 1A, middle right) and TorA (KK) (Fig. 1A, bottom right). In these mutants, the twin arginine residues in the signal sequence were replaced with twin lysine residues, resulting in an inability to translocate across the IM. We observed that these KK mutants did not convert to the mature form, even in TatABC-overproducing strains (Fig. 1A, middle). These results indicate that the TAT translocation was properly studied. We investigated the involvement of MPIase in more detail using pulse-chase experiments (Fig. 1B). To facilitate TAT translocation, TatABC was overexpressed from the tac promoter on a plasmid (pTac-TatABC). In the wild-type strain harboring pTac-TatABC (Fig. 1B, top left), the precursor form of SufI was efficiently converted to the mature form. No mature form was observed when the signal mutant SufI (KK) was used (Fig. 1B, top right), confirming that the TAT translocation was analyzed correctly. When MPIase was depleted, no mature form was observed even after a 10-min chase (Fig. 1B, bottom, ‘ΔMPIase’), confirming MPIase’s involvement in the TAT system. When MPIase synthesis was induced (Fig. 1B, bottom, ‘+MPIase’), the mature form appeared, but the conversion rate was much slower than under wild-type conditions. The reason for this will be discussed later. To investigate the role of MPIase in the TAT system, we performed cell fractionation into periplasmic and spheroplast fractions (Fig. 1C). When the MPIase-depleted strain expressing TatABC and SufI was fractionated, the precursor was recovered in the spheroplast fraction, and no mature form was observed. This indicates that SufI was not translocated (Fig. 1C, left, ‘ΔMPIase’). Conversely, mature SufI was translocated to the periplasm, when MPIase was expressed (Fig. 1C, left, ‘+MPIase’). Successful fractionation was confirmed by detecting the periplasmic protein β-lactamase and the cytoplasmic protein SecB. Similar results were obtained when TorA-GFP, in which the TorA signal is connected to GFP, was examined (Fig. 1C, right), confirming that TorA-GFP is not translocated upon MPIase depletion. These cells expressing TorA-GFP were also used to detect the subcellular localization directly by monitoring its fluorescence (Fig. 1D). In the presence of MPIase, the envelope fraction exhibited green fluorescence, indicating that GFP is localized in the periplasm or on the inner membrane at least (left). Conversely, in the absence of MPIase, the entire cell exhibited uniform fluorescence, indicating that TorA-GFP is localized in the cytoplasm under these conditions (right). These results are consistent with those of the subcellular fractionation (see Fig. 1C). Taken together, we conclude that MPIase is involved in the TAT system in vivo.

MPIase depletion causes the accumulation of TAT substrate precursors.

(A) Detection of SufI and TorA precursors upon MPIase depletion. SufI (top) and TorA (bottom) were expressed in the indicated strains. IPTG (1.5 mM) was added at early log phase, and cultivation was continued for 2 h. The cells were subjected to TCA precipitation, followed by SDS-PAGE/immunoblotting. The precursor and mature forms were then detected as indicated. MPIase depletion was confirmed in KS23 (middle left). Defects in the processing of SufI (middle right) and TorA (bottom right) with the KK signal were also confirmed. The symbol ‘★’ denotes degradation products of the SufI precursor. (B) Pulse-chase experiments of SufI processing after MPIase depletion. EK413/pTac-TatABC harboring pTet-SufI(RR) (top left) or pTet-SufI(KK) (top right), and KS23/pAra-CdsA/pTac-TatABC harboring pTet-SufI(RR) (+MPIase; bottom left, ΔMPIase; bottom right) were pulse-labeled for 30 s and then chased for the indicated periods. SufI(RR) and SufI(KK) were isolated using the attached His tags, followed by SDS-PAGE/autoradiography. (C) Subcellular localization of SufI and TorA-GFP after MPIase depletion. KS23/pTet-TatABC/pTac-SufI(RR) (left) and KS23/pTet-TatABC/pTac-TorA-GFP(RR) (right) were cultivated in the presence (+MPIase) or absence (ΔMPIase) of arabinose. The TAT substrates were leaky expressed from the tac promoter. The cells were then fractionated into periplasmic (peri) and spheroplast (sph) fractions. The indicated proteins were detected by immunoblotting. (D) Subcellular localization of TorA-GFP after MPIase depletion as revealed by fluorescence microscopy. The indicated strains were cultivated, TorA-GFP was induced, and then it was observed by fluorescence microscopy. The scale bars represent 10 μm. A magnified view of part of the ‘+MPIase’ sample is shown on the left.

TAT substrates are translocated into INV in vitro when TatABC and MPIase are both overproduced.

(A) Translocation of SufI and TorA-GFP into TatABC-overproducing INV. SufI (top) and TorA-GFP (bottom) were synthesized in vitro in the presence of INV prepared from the indicated strains. INV was omitted in the ‘-’ samples. The signal sequences on the left are wild-type (RR), while the sequences on the right are mutants (KK). Translocated materials were generated after PK digestion (+). One-fifth of the reaction mixture was analyzed as the translation standard (-). Translocation activities, expressed as a percentage of the amount of the synthesized proteins, are shown at the bottom of each autoradiogram. (B) Depletion of MPIase resulted in the failure of TAT translocation. INV were prepared from KS46/pAra-CdsA harboring pT7-TatABC, which was cultivated in the presence (+MPIase) or absence (ΔMPIase) of arabinose. After washing arabinose, IPTG was added at 1.5 mM to induce TatABC, and cultivation was continued for 2 h prior to INV preparation. SufI translocation was then performed as described in (A). (C) The level of TatC in INV used in (A) and (B). Each INV (10 μg) was analyzed by SDS-PAGE, followed by immunoblotting using anti-RpmJ-His antibodies. The position of TatC-His was shown. (D) MPIase is upregulated upon TatABC overproduction. The expression levels of MPIase in the strains used in (B) were determined by immunoblotting. CdsA and TatABC were induced as indicated. SecB levels were determined as a loading control. MPIase is co-purified with TatABC. INV prepared from BL21/pT7-TatABC-His induced with 1.5 mM IPTG (+++) or not induced (-), and BL21/pT7-Pm-F0b-His induced with 1 mM IPTG, solubilized with DDM, were applied to a TALON column, then eluted. The eluates were analyzed by TLC, followed by immunodetection of MPIase. The positions of the origin/front and MPIase are shown. Purified MPIase was also analyzed at right.

Upregulation of MPIase upon TatABC overproduction is essential for in vitro Tat translocation

The TAT pathway can be reproduced in vitro using inverted membrane vesicles (INV), as depicted in Fig. S3. However, it has been shown that overproduction of TatABC is necessary for translocation into INV 10,59. To reproduce these observations, we prepared INV from wild-type strains with an empty vector (Fig. 2A BL21/pT7 (empty)) and from TatABC-overproducing strains (Fig. 2A, BL21/pT7-TatABC), which were then subjected to the SufI translocation assay in vitro. TatABC overproduction with this vector was confirmed by immunoblotting (Fig. S4). As reported 10,59, SufI translocation into INV was observed only when TatABC was overproduced (Fig. 2A, top left). In the absence of INV, SufI was completely digested by PK (-). These results indicate that SufI is translocated in a TatABC-dependent manner. SufI (KK) was not translocated, even when TatABC-overproducing INV were used (Fig. 2A, top right), indicating that the TAT pathway is adequately analyzed in our experimental system. We observed the same trend with another TAT substrate, TorA-GFP (Fig. 2A, bottom). To further test the involvement of MPIase in the TAT pathway in this in vitro system, we used a cdsA/ynbB knockout strain with a BL21 (DE3) background (KS46) harboring pAra-CdsA. We then prepared MPIase-depleted INV with TatABC overproduction, followed by SufI translocation (Fig. 2B). Since TatABC was overproduced prior to MPIase depletion, these INV contain the overproduced amount of TatABC (Fig. 2C) but not MPIase (Fig. 2D). We observed that SufI was not translocated into these MPIase-depleted INV, which further supports the involvement of MPIase. Consistently, the anti-MPIase antibody inhibited SufI translocation (Fig. S5). However, we also found that the translocation activity with MPIase-induced INV (Fig. 2B, ‘+MPIase’) did not recover to the level observed with the wild-type INV with TatABC overproduction (Fig. 2B, ‘BL21/pT7-TatABC’). The reason for the incomplete recovery was that TatABC overproduction caused MPIase upregulation (Fig. 2D). Significant MPIase upregulation occurred when TatABC overproduction was induced in BL21(DE3). However, MPIase expression remained unchanged with TatABC overproduction in KS46, in which cdsA was induced from the plasmid pAra-CdsA (Fig. 2D). Since the levels of Tat components in BL21(DE3) and KS46/pAra-CdsA grown with arabinose were similar (Fig. 2C), it is reasonable to conclude that MPIase upregulation is necessary for the translocation activity in vitro of the TAT pathway into INV. This explains the low activity of MPIase-induced INV (Fig. 2B, ‘KS46/pAra-CdsA +MPIase’). These results are consistent with the fact that the SufI translocation was not fully restored in the pulse-chase experiments when MPIase was induced from pAra-CdsA (see Fig. 1B). MPIase upregulation was specific to TatABC overproduction; MPIase levels were barely affected by SecYEG overproduction 52 (Fig. S6). These results strongly suggest that TatABC and MPIase function cooperatively. To examine whether the interaction between TatABC and MPIase is direct and physical, we detected MPIase in the TatABC-His preparation purified via metal affinity chromatography. As shown in Fig. 2E, MPIase was co-purified with TatABC. Furthermore, the amount of MPIase precipitated with TatABC increased significantly with TatABC overproduction (Fig. 2E). When His-tagged F0b from Propionigenium modestum (Pm-F0b-His) was expressed and used as a control, no detectable amount of MPIase was precipitated. Under these conditions, both TatABC-His and Pm-F0b-His were purified to homogeneity (Fig. S7). These results indicate that TatABC and MPIase directly and specifically interact with each other. Based on these results, we concluded that the TatABC requires MPIase to function.

MPIase is the missing piece for successfully reconstituting the TAT pathway

Results obtained thus far, both in vivo and in vitro, demonstrate the essentiality of both TatABC and MPIase. Therefore, we assumed that the TAT reconstitution might be possible if TatABC and MPIase were both present and PMF was imposed. To test if our reconstitution system would work, we started with a crude system using proteoliposomes reconstituted from the solubilized membranes. First, we prepared proteoliposomes from solubilized membranes with overproduced amounts of TatABC and purified F0F1-ATPase from thermophilic bacteria (see Fig. 3B, left), followed by inclusion of DAG for stable generation of PMF 61. However, these proteoliposomes were inactive with regard to SufI translocation (Fig. 3A, top left, ‘+PL/DAG’). We found that fusing MPIase-containing liposomes to the proteoliposomes greatly enhanced SufI translocation (Fig. 3A, top left, ‘+PL/DAG/MPIase’). This may be because MPIase was not efficiently recovered in the proteoliposome fraction due to the hydrophilic nature of MPIase’s glycan moiety 46. These results again indicate that MPIase is essential for the TAT system. Even when MPIase-containing liposomes were fused, SufI (KK) was not translocated into the proteoliposomes (Fig. 3A, bottom left). We also examined the post-translational translocation using a partially purified SufI precursor (Fig. 3A, top right). We observed SufI (RR) translocation into TatABC/MPIase proteoliposomes in an ATP-dependent manner (i.e., a PMF-dependent manner), but not SufI (KK) translocation. Thus, the TAT reconstitution was achieved through a combination of TatABC, MPIase and F0F1-ATPase.

Reconstitution of the TAT system.

(A) Membrane fusion of TatABC proteoliposomes with MPIase liposomes enabled the detection of the SufI translocation. INV, prepared from BL21/pT7-TatABC, were solubilized with OG, and then mixed with purified F0F1-ATPase. The proteoliposomes were then formed by removing the detergent. PL/DAG liposomes or PL/DAG/MPIase liposomes were fused to the reconstituted proteoliposomes. SufI translocation was performed co-translationally (left) or post-translationally (right) in the presence of these proteoliposomes. SufI(RR) (top left) or SufI(KK) (bottom left) was used as the substrate. One-ninth of the reaction mixture was analyzed as the translation standard (-) for the co-translational reactions. For the post-translational reactions, ATP was added as indicated. Translocation activities are shown at the bottom of the left panels. (B) Purification of F0F1-ATPase and the Tat components. F0F1-ATPase from a thermophilic Bacillus PS3 (left), and E. coli TatA-His (lane 1), TatB-His (lane 2), and TatC-His (lane 3) (right) were subjected to SDS-PAGE, followed by Coomassie Brilliant Blue staining. (C, D) Complete reconstitution of the TAT system using the purified components. (C) Purified TatA, TatB, and TatC (in a 1:1:1 ratio) and F0F1-ATPase were mixed with phospholipids. Then, detergent was removed to form proteoliposomes. The reconstituted proteoliposomes were then fused with PL/DAG liposomes or PL/DAG/MPIase liposomes. Translocation activity was assayed with SufI(RR) (left) or SufI(KK) (right). (D) Proteoliposomes were reconstituted with the indicated components, followed by TorA-GFP translocation. The TatA:TatB:TatC ratio was 10:1:1. CCCP (20 μM), a protonophore, was added to the sample on the left. TorA-GFP(RR) (top) or TorA-GFP(KK) (bottom) was used as the substrate. Translocation activities are shown at the bottom of each gel.

Next, we purified TatA, TatB and TatC individually through an attached His tag (Fig. 3B, right). We confirmed that expressing each His-tagged TatA, TatB, and TatC complemented the SDS-sensitive phenotype of the corresponding knockout strain (Fig. S8). Using these purified proteins, we reconstituted proteoliposomes containing TatABC and F0F1-ATPase. Then, we fused them with MPIase/DAG liposomes. The TatABC-MPIase proteoliposomes containing F0F1-ATPase exhibited activity with respect to SufI translocation (Fig. 3C, left). This activity was comparable to that of proteoliposomes reconstituted with solubilized membranes (Fig. 3A). On the other hand, SufI (KK) was not translocated into the proteoliposomes (right). Similar results were obtained when using TorA-GFP (Fig. 3D). Proteoliposomes containing both TatABC and MPIase allowed for the translocation of TorA-GFP (RR) but not TorA-GFP (KK). The translocation of TorA-GFP (RR) was severely inhibited by CCCP even if the reaction mixture contains DTT that inactivates CCCP, indicating that the translocation is dependent on PMF (right). Furthermore, omitting MPIase, TatABC or F0F1-ATPase resulted in a loss of translocation activity (left). These results indicate that TatABC, MPIase and PMF are necessary and sufficient for TAT translocation.

MPIase serves as the TAT signal receptor

Although the participation of MPIase in the Tat pathway has been demonstrated both in vivo and in vitro, the mechanisms by which MPIase contributes to Tat translocation remains unclear. Our previous study showed that the glycan chain of MPIase captures the substrate membrane protein on the membrane surface before membrane insertion 48,56. Although the TatBC complex has been shown to act as the TAT signal sequence receptor 26,30, we hypothesized that MPIase could also be involved in the targeting of the TAT substrate proteins to the membrane. Thus, we examined whether or not MPIase is involved in the membrane localization of TorA-GFP. Cells expressing TorA-GFP were fractionated into soluble (cytosol plus periplasm) and membrane fractions. Then, we measured the GFP fluorescence. We confirmed that the cell debris and microaggregates could be removed by low-speed centrifugation before membrane sedimentation (Fig. S2). When we induced MPIase in KS23 cells by adding arabinose, ∼30% of the fluorescence was recovered in the membrane fraction (Fig. 4A, ‘RR, +MPIase’), similarly to the ‘EK413’ case. Note that the MPIase level was insufficient under these conditions, resulting in very low translocation efficiency (see Fig. 1B). Conversely, when MPIase was depleted, only <10% of the fluorescence was recovered in the membrane fraction (Fig. 4A, ‘RR, ΔMPIase’). Similar results were obtained using TorA-GFP with a mutated signal sequence (KK) (Fig. 4A). These results strongly suggest that MPIase serves as a receptor for the TAT signal, but that MPIase does not discriminate the functional signal. Similar to the ‘+MPIase’ and ‘EK413’ cases, ∼30% of the fluorescence was recovered in the membrane fraction when the tatC knockout mutant was used (Fig. 4A, ‘ΔTatC’). These results suggest that MPIase recognizes the TAT signal before TatC does. Since TatABC overproduction causes MPIase overproduction (see Fig. 2D), we examined the effect of TatABC/MPIase overproduction on TorA-GFP localization (Fig. 4B). As expected, nearly 90% of TorA-GFP was recovered in the membrane fraction when EK413 (cdsA+) cells, in which both TatABC and MPIase were overproduced, were used (Fig. 4B, ‘RR, EK413’). In marked contrast, when MPIase was depleted, only 10∼20% of TorA-GFP was recovered in the membrane fraction, despite TatABC overproduction (Fig. 4B, ‘RR, ΔMPIase’). These results confirm that MPIase serves as a receptor for the TAT signal. When KS23 cells with TatABC overproduction were cultivated in the presence of arabinose to induce MPIase, the amount of membrane-targeted TorA-GFP increased significantly (∼50%; Fig. 4B, ‘RR, +MPIase’) compared to the absence of TatABC overproduction (∼30%; Fig. 4A), indicating that TatABC stabilizes membrane-bound TAT precursors, consistent with previous reports 59. Nonetheless, this level was much lower than that in the cdsA+ conditions (Fig. 4B, ‘RR, EK413’), in which MPIase is upregulated by TatABC overproduction. These results again indicate that MPIase serves as a TAT signal receptor. Similar results were obtained when using the KK mutant of TorA-GFP (Fig. 4B, right half), indicating that MPIase cannot discriminate against the functional signal sequence. However, the level in the cdsA+ conditions (Fig. 4B, ‘KK, EK413’) was lower than that with TorA-GFP (RR) (Fig. 4B, ‘RR, EK413’), suggesting that TatABC excluded the KK mutant. When GFP without the TAT signal was expressed, >90% of fluorescence was recovered in the supernatant fraction, regardless of MPIase levels and TatABC overproduction (Fig. S9), indicating that the membrane localization depends solely on the TorA signal.

MPIase is a membrane receptor for TAT substrates.

(A-C) TorA-GFP is targeted to the membrane in an MPIase-dependent manner. TorA-GFP(RR) (A, B; left half) or TorA-GFP(KK) (A, B; right half) was expressed in the specified strains. The cells were then disrupted to obtain the supernatant (sup; cytosol and periplasm) and the precipitate (ppt; membranes) fractions. The fluorescence intensity of each fraction was measured. The ratio to the whole cell was determined and shown as percentages. TatABC was overexpressed in (B), but not in (A). The means and standard deviations were calculated from at least three independent experiments. *, P<0.05; **, P<0.01. In (C), the indicated strains were cultivated, TorA-GFP was induced, and then it was observed by fluorescence microscopy. The scale bars represent 10 μm. Part of each picture was magnified on each side. Scale bars represent 10 μm. (D) TorA-GFP interacts with INV expressing MPIase. Purified TorA-GFP was mixed with the indicated INV, followed by fluorescence detection (top). As a control, each INV was mixed with NBD-PE, a fluorescent phospholipid, and observed (bottom). Scale bars represent 10 μm.

The MPIase-dependent membrane targeting was also observed using fluorescent microscopy (Fig. 4C). When MPIase was induced by adding arabinose, an envelope-shaped fluorescence signal was detected, regardless of whether it was wild-type or mutant (Fig. 4C, left), indicating that TorA-GFP (RR) and TorA-GFP (KK) were both targeted to membranes. Conversely, under the MPIase-depleted conditions, whole-cell fluorescence was detected (Fig. 4C, right), indicating that both TorA-GFP (RR) and TorA-GFP (KK) located in the cytosol.

MPIase-dependent targeting was confirmed in vitro (Fig. 4D, top). When the purified TorA-GFP (Fig. S10) was mixed with wild-type INV (Fig. 4D, ‘EK413’), TatABC-overproducing INV (Fig. 4D, ‘TatABC+++’) and TatC-deficient INV (Fig. 4D, ‘ΔTatC’), in all of which contain MPIase, fluorescent dots were observed, indicating that TorA-GFP directly interacts with these INV. The number of dots increased with TatABC overproduction, which is consistent with MPIase upregulation induced by TatABC overproduction. In contrast, no such fluorescent dots were observed in the presence of MPIase-depleted INV (Fig. 4D, ‘ΔMPIase’), confirming the MPIase-dependent targeting of TorA-GFP. The presence of similar amounts of membrane vesicles was visualized by adding fluorescent phospholipids, NBD-PE (Fig. 4D, bottom).

Discussion

In this study, we demonstrated that glycolipid MPIase is essential for the TAT system. All of the in vivo, in vitro and reconstitution studies supported this conclusion. Therefore, the previous unsuccessful attempts at TAT reconstitution were likely due to an insufficient amount of MPIase in TatABC proteoliposomes. Even when using solubilized membranes containing TatABC and MPIase to prepare proteoliposomes, MPIase recovery would be too low to detect the TAT activity.

MPIase is a glycolipid that catalyzes the membrane insertion of proteins with a single TM 50,53 as well as those with multiple TM domains 51,55,57. Therefore, depleting MPIase should impair the membrane insertion of all the TAT components, TatA, TatB and TatC, leading to the failure of TAT translocation. However, in our experimental system, TatABC was still present after MPIase depletion (see Fig. 2C), which supports the direct involvement of MPIase in the TAT translocation. Previous studies have reported that the Sec translocation slows down but does not stop completely, when MPIase is depleted 58, suggesting that the Sec machinery, including SecYEG, remains present during the initial stages of MPIase depletion, as demonstrated in this study. The results of the reconstitution study, in which TatABC was reconstituted in proteoliposome membranes, rule out the indirect involvement of MPIase in TAT translocation. In this system, the TAT translocation occurred in TatABC-, PMF- and MPIase-dependent manners.

How is then MPIase involved in the TAT system? Since MPIase has been identified as a factor that catalyzes membrane protein integration 46,50, it is possible that MPIase drives the insertion of the TAT signal sequences into membranes, thereby initiating the translocation reaction (Fig. 5, (i)). Depleting MPIase caused TorA-GFP to accumulate in the cytosol. Thus, MPIase serves as a receptor for TAT precursors, and the TAT signal sequences are inserted into membranes. Then, they are transferred to TatC (ii), where the functional TAT signal is verified (iii). In addition to its roles as a membrane receptor and an insertase for the TAT signals, MPIase may facilitate structural changes in TatABC, allowing the complex to form the translocon dedicated to the substrate protein (iv). The TAT substrate is then translocated through the translocon, and the mature protein is released into the periplasm upon Lep-mediated cleavage of the signal sequence 10,43 (v). These processes require PMF, as previously reported 62. Thus, MPIase may be involved not only in the membrane targeting of TAT substrates, but also in forming the TatABC translocon for each substrate. Membrane lipids PE may be important for these processes, since PE is essential for TAT translocation 63. In the reconstitution studies, the absence of DAG in (proteo)liposomes allowed for the spontaneous insertion of membrane proteins and signal sequences 55,64. Therefore, the failure of the TAT reconstitution strongly suggests that MPIase has functions beyond serving as a membrane receptor and an insertase for TAT translocation. Given that MPIase induces dynamic structural changes in the SecYEG translocon 52,54, it is likely that MPIase also modifies the structure of TatABC to form a tailor-made translocon. This possibility is consistent with the observations showing that TatABC overproduction induces MPIase upregulation and that MPIase co-precipitates with TatABC.

Working model of mechanisms underlying MPIase-dependent TAT translocation.

MPIase recognizes the signal sequence of the TAT substrates and insert them into the membrane. The TatBC complex on the membrane then determines whether the signal sequence of the membrane-inserted TAT substrate is valid. The TatABC translocon then forms a pore the same size as the substrate. This allows the substrate protein to be translocated across the membrane. Finally, Lep cleaves the signal sequence, and the mature protein is released into the periplasm. See the text for details.

It is well known that the lipid environment of the membranes significantly affects the efficiency of TAT translocation. First, acidic phospholipids are important for TAT translocation because TAT signals interact with the acidic phospholipids 65. MPIase has a pyrophosphate residue and would therefore contribute to this interaction in a similar way. At the same time, the hydrophobic domain of the TAT signals is solubilized by the glycan chain, as occurs with membrane proteins 47. Conversely, a decrease in PE (i.e., an increase in acidic phospholipids, PG and CL) decreases efficiency 66,67, presumably due to an overly strong interactions between the TAT signals and the acidic phospholipids. Alternatively, PE may be important for TAT translocation, since it stabilizes several membrane proteins 68. Thus, PE may be involved in the expression of the TatABC function. In the thylakoid system, the reduced unsaturation in membrane lipids results in a 50% decrease in the efficiency of TAT translocation 69, suggesting that membrane fluidity is also important for TAT translocation. We have demonstrated that MPIase increases the membrane fluidity through the flexible movement of its long sugar chain 49. This property would benefit TAT translocation.

MPIase is upregulated in the cold. During this process, both cdsA and ynbB are induced at the transcriptional level. In the case of cdsA induction, two cold-inducible promoters located upstream of the cdsA gene are responsible 70. MucA, a small peptide that overlaps with cdsA, enhances MPIase synthesis by CdsA 71. When MPIase is upregulated by TatABC overproduction (see Fig. 2D), cdsA and/or ynbB are also induced, as evidenced by the diminished increase in the MPIase level in cdsA/ynbB knockouts even when CdsA expression is complemented from the araBAD promoter on a plasmid. In this case, the cdsA promoters other than identified cold-inducible ones would be responsible, together with MucA. Alternatively, other mechanisms could operate to upregulate MPIase.

We found that the TAT translocation could be reconstituted with TatABC and MPIase alone when PMF was imposed, indicating that these factors are sufficient for TAT translocation. Although the activity was comparable to that of INV and proteoliposomes with the crude extract, it was still low. Therefore, these results suggest the presence of factor(s) involved in TAT translocation that stimulate the reaction, as the soluble factors that solubilize the TAT substrates and target them to membranes remain unidentified.

The TAT system is widely used as a carrier system for secreting foreign proteins into the medium 72–74. Now that we know that MPIase is involved in the TAT system, it will be easier to engineer the TAT system by modulating the MPIase activity to increase the level of secreted proteins.

The TAT system was first identified in the thylakoid membranes of chloroplasts as a ΔpH-dependent pathway 2,75. The components of the TAT system in chloroplasts are highly homologous to TatABC in E. coli 26,37, and the translocation reaction is also driven by PMF 2. Thus, the mechanisms underlying the TAT systems in E. coli and chloroplasts likely have similar underlying mechanisms, strongly suggesting the presence of an MPIase homologue in chloroplasts. This similarity could facilitate the identification of an MPIase homologue in chloroplasts.

Methods

Materials

The strains, plasmids, and oligonucleotides used in this study are listed in Tables S1, S2, and S3, respectively. INV were prepared as described 59. The plasmid pCspB-TorAss-GFP encoding TorA-GFP was obtained from Dr. Y. Kikuchi (Ajinomoto Co., Inc.). Bio Beads SM-2 were purchased from BioRad. [35S] EXPRE35S35S Protein Labeling Mix, a mixture of [35S] methionine and [35S] cysteine (∼37 TBq/mmol), was obtained from Revvity. Proteinase K (PK) was purchased from Roche Diagnostics. The cell-free protein synthesis kit “E. coli S30 Extract System for Circular DNA” was purchased from Promega Corporation. The E. coli polar lipid extract (phospholipids), 18:1 NBD-PE, and dioleoylglycerol were obtained from Avanti Polar Lipids, Inc. n-Octyl-β-D-glucopyranoside (OG) and n-dodecyl-β-D-maltopyranoside (DDM) were obtained from Dojindo Laboratories. IPTG was from Wako Pure Chemicals Industries, Ltd. TLC plates were from Merck. MPIase was purified from MC4100, as described 46,50. TorA-GFP, TatA, TatB, TatC, and the TatABC complex were purified from DH5α/pTac-TorA-GFP-His, DH5α/pTac-TatA-His, DH5α/pTac-TatB-His, DH5α/pTac-TatC-His, BL21/pT7-Pm-F0b-His, and BL21/pT7-TatABC-His, respectively. These proteins were induced with 1.5HN and JSPS KAKENHI grants mM IPTG, followed by INV preparation, solubilization and purification through the 6 x His-tag attached to C-termini on Talon columns. The conditions for INV solubilization and Talon column chromatography was carried out as described 57. Purification of F0F1-ATPase from a thermophilic bacterium PS3 was performed as described 76.

Depletion of MPIase

Strains KS23/pAra-CdsA and KS46/pAra-CdsA were cultivated overnight at 37°C in LB medium containing 0.002% arabinose. After the cells were washed five times with fresh LB medium, they were inoculated into LB medium supplemented or not with 0.2% arabinose at a dilution of 1/1000 and then shaken at 37°C. To express the TAT substrates, IPTG (1 mM) was added after the cell culture was grown to OD600 nm ∼0.5 and cultivated for another 1 h. An aliquot (500 µL) of each sample was treated with TCA to a final concentration of 5%. The TCA precipitate was recovered by centrifugation (10,000 x g, 4°C, 5 min), followed by washing with acetone. MPIase depletion of was then confirmed by SDS-PAGE/immunoblotting, as described58.

Pulse-chase experiments

Cell cultures grown in LB medium to OD600 nm ∼0.5 were harvested and resuspended in M9 minimal medium containing 18 amino acids (0.1 mM each) except for methionine and cysteine, and 0.1% yeast extract. After incubating for an additional 1 h at 37°C, the substrate proteins were labeled with [35S] methionine/cysteine (∼10 MBq/mL) for 30 s. Labeling was terminated by adding non-radioactive methionine and cysteine to a final concentration of 12 mM. At the specified chase time, an aliquot (500 µL) of the culture was withdrawn, and TCA was added to a final concentration of 10%. The TCA precipitates were washed successively with acetone and diethyl ether, and then suspended in 50 µL of 50 mM Tris-HCl (pH 7.5), 1% SDS, 1 mM EDTA (pH 8.0), and boiled for 3 min. The samples were then diluted with 1 mL of 50 mM HEPES-KOH (pH 7.5), 150 mM NaCl, 1% Triton X-100, 1 mM EDTA (pH 8.0), 1 mM phenylmethylsulfonyl fluoride, and were centrifuged at 16,000 x g for 5 min. Then, 10 µL of TALON resin was added to the supernatant, and the mixture was stirred at 4°C for 1 h. After centrifugation (3,000 x g, 4°C, 30 s), the TALON resin was washed successively with 500 µL of 50 mM Tris-HCl (pH 7.5), 150 mM NaCl, 1% Triton X-100. The washed resin was suspended in 20 µL of the sample buffer for SDS-PAGE. Radioactive bands, analyzed by SDS-PAGE, were detected by autoradiography using a phosphorimager (GE Healthcare), and quantified using ImageQuant (GE Healthcare).

Preparation and fractionation of spheroplasts

After the cells were grown in LB medium to OD600 nm ∼0.5, 1 mM IPTG was added to induce the TAT substrates. The cells were harvested at 2,400 x g for 5 min and washed with 5 mL of 0.75 M sucrose, 10 mM Tris-HCl (pH 7.5). They were then suspended in 1 mL of 0.75 M sucrose, 10 mM Tris-HCl (pH 7.5) and 20 µg/mL lysozyme solution and incubated on ice for 5 min. They were converted into spheroplasts by adding 1.5 mM EDTA (pH 8.0) at equal intervals over 10 min, bringing the total volume to 2 mL. An aliquot of the spheroplast suspension (1 mL) was centrifuged (15,000 x g, 4°C, 1 min) to sediment spheroplasts. The spheroplasts were suspended in 500 µL water and treated with 10% TCA, followed by incubation on ice for 10 min to precipitate proteins. The periplasmic fraction, obtained by sedimentation of the spheroplasts, was treated similarly. The proteins were then recovered by centrifugation (10,000 x g, 4°C, 5 min), washed with acetone, and subjected to SDS-PAGE/immunoblotting, as previously described 77.

Fluorescence microscopy

E. coli strains were grown in LB medium to OD600 nm ∼0.5, followed by the addition of 1 mM IPTG to induce TorA-GFP. The cells were fixed on glass slides coated with 1% agarose, as described 39. The cells were observed using a fluorescence microscope (Olympus BX53, combined with a U-FGFP filter). Images were captured with a 3CCD digital camera (Olympus DP74). The digital images were then processed and analyzed using the cell image analysis software, cellSens Standard (Olympus).

Assaying of the protein translocation in vitro

The TAT substrates were synthesized in vitro using the E. coli T7 S30 extract system. For the co-translational translocation assay, the followings were added to a 20 µL reaction mixture: [35S] methionine and cysteine (∼10 MBq/mL), plasmids encoding the substrate protein under the T7 promoter, and INV or proteoliposomes. The synthesis/translocation of the RI-labeled substrates were performed at 37°C for 1 h. When INV was used, it was added to the reaction mixture 5 min after the reaction started. For the post-translational translocation assay, the RI-labeled substrates (3 µL), synthesized in the absence of INV or proteoliposomes at 30°C for 30 min, were added to a translocation mixture containing 50 mM HEPES-KOH (pH 7.5), INV or proteoliposomes, 1 mM MgSO4, 5 mM creatine phosphate, 10 µg/mL creatine kinase, and 1 mM ATP. After incubating the mixture at 37°C for 45 min, it was divided into two portions. One portion was treated with 5% TCA to precipitate the substrate proteins. The other portion was mixed with an equal volume of PK (1 mg/mL) and incubated at 25°C for 20 min. After adding 5% TCA, the samples were incubated at 56°C for 5 min to inactivate PK. Then, the proteins were recovered by centrifugation (10,000 x g, 4°C, 5 min). Radioactive materials were detected and quantified as described above.

TALON affinity chromatography

INV (4 mg protein) was solubilized using 1% DDM, 20% glycerol, 50 mM HEPES-KOH (pH 7.5). The solubilized membranes were recovered by centrifugation (170,000 x g, 4°C, 30 min). Samples were applied to a TALON column (0.5 mL). After washing the column with 20 mL of 0.02% DDM, 10% glycerol, and 50 mM HEPES-KOH (pH 7.5), the column-bound materials were eluted with the same buffer containing 150 mM imidazole. The eluted fraction was analyzed by TLC using a solvent system C (chloroform/ethanol/water: 3/7/4) for development. Then, MPIase was detected using anti-MPIase antibody, as described 53.

Reconstitution of proteoliposomes

Phospholipids (13 mg) were resuspended in buffer A (50 mM HEPES-KOH (pH 7.5), 1 mM dithiothreitol) at 20 mg/mL, followed by bath sonication to form liposomes. The liposomes were solubilized in 5% [w/v] OG and incubated on ice for 15 min. Then, 200 mg of Bio Beads SM-2 were added, and the mixture was gently stirred at 4°C for 30 min. Purified TatA (100 µg or 1000 µg), TatB (100 µg), TatC (100 µg) and F0F1-ATPase (1 µM) were added to the solution. The solution was gently mixed at 4°C for at least 3 h. The Bio Beads SM-2 were then removed, and the proteoliposomes were recovered by centrifugation (170,000 x g, 4°C, 1 h). DAG- and DAG/MPIase-containing liposomes were prepared, as described 77. TatABC/F0F1-ATPase proteoliposomes and PL/DAG liposomes or PL/DAG/MPIase liposomes were typically mixed at a ratio of 4:1, followed by three cycles of freezing, thawing, and sonicating to allow liposome fusion, as described 54. The mixture was then passed through a mini-extruder (Avanti Polar Lipids, Inc.) with a 0.4 µm pore size filter to obtain uniformly sized liposomes, as described 64.

Subcellular fractionation

After the cells were grown in LB medium to OD600 nm ∼0.5, 1 mM IPTG was added to induce the TAT substrates. The harvested cells were suspended in 50 mM HEPES-KOH (pH 7.5), 10% glycerol, 1 mM dithiothreitol, 1 mM phenylmethylsulfonyl fluoride, protease inhibitor cocktail (Roche) and then disrupted using a French press. Then, the cells were incubated with 2 mM MgSO4, DNase I (1 mg/mL) on ice for 15 min, followed by the addition of 2 mM EDTA (pH 8.0) and incubation on ice for an additional 15 min. The disrupted and undisrupted cells were separated by centrifugation (15,000 x g, 4°C, 10 min) and then they were fractionated into soluble (cytosol and periplasm) and insoluble (membrane) fractions by centrifugation (150,000 x g, 4°C, 1 h). TorA-GFP was detected by monitoring GFP fluorescence. Samples were added to 2 mL of 50 mM HEPES-KOH (pH 7.5), and the fluorescence intensity was measured at an excitation of 395 nm and an emission of 509 nm using an FP-8200 fluorescence spectrometer (JASCO).

Statistical analysis

Statistical significance was determined using a Student’s t-test. The P values of less than 0.05 were considered significant.

Other methods

SDS-PAGE was performed using 12.5% acrylamide/0.27% N,N’-bis (methyleneacrylamide) 78 for TatA, TatB, TatC, SecB, β-lactamase, and MPIase. 15.0% acrylamide/0.33% N,N’-bis (methyleneacrylamide) 79 was used to analyze SufI and TorA-GFP, while 8% acrylamide/0.17% N,N’-bis (methyleneacrylamide) was used for TorA. Bands on the immunoblots were quantified using a CS analyzer (ATTO). Protein quantification after INV preparation was performed using bovine serum albumin as a standard, as described 80. Detection of MPIase and other proteins by immunoblotting was performed as described 58,81.

eLife Assessment

This study demonstrates a critical role of the glycolipid membrane protein insertase (MPIase) in the twin-arginine translocation (Tat) pathway, a protein translocation system that is conserved across all domains of life. The successful reconstitution of the bacterial Tat system in both bacteria-derived and artificial liposomes provides solid experimental evidence that MPIase has a broader role in membrane protein translocation than previously recognized. By revealing the essential function of a nonproteinaceous membrane component in catalyzing a core cellular process, this work offers fundamental insights into the molecular mechanisms of protein translocation.

Reviewer #1 (Public review):

Hanako and colleagues demonstrated that glycolipid MPIase is essential for the TAT system, and they successfully reconstituted the TAT system in vitro for the first time. This will facilitate the understanding of the mechanism of the TAT system.

My major points are listed below for the authors to consider:

(1) The authors successfully reconstituted the TAT system using the purified TatA/B/C, but the translocation efficiency was much lower than that of native INV. The authors partly attributed this to the reason that "MPIase recovery would be too low to detect the TAT activity" in the Discussion part. So, what would happen to the translocation efficiency if you added more MPIase to the reconstituted system? How about the abundance of MPIase from the INV and reconstituted proteoliposomes?

(2) Why were only TatC levels measured in Figure 2C, whereas the expression levels of TatA were not detected? Also, from my observation, the amount of TatC in the third lane is lower than that in the previous two lanes.

(3) The authors should explain why the TatA/B/C ratios in Figure 3C (1:1:1) and Figure 3D (10:1:1) are inconsistent.

(4) ~30% of the fluorescence was recovered in the membrane fraction (Figure 4A) both in the functional TAT signal sequence (RR) and in the inactivating mutant signal sequence (KK), which suggests that MPIase acts as a relatively broad recognition factor. Given that MPIase does not discriminate between RR and KK, why do un-translocated substrates remain in the cytoplasm rather than non-specifically adhering to the membrane when MPIase is depleted in vivo?

Reviewer #2 (Public review):

Summary:

In this manuscript, the authors investigated the relationship between the Tat system and MPIase, a glycolipid that facilitates protein integration into the bacterial cell membrane. The TAT (twin-arginine translocation) system is a unique membrane transport machinery that exports fully folded proteins containing a twin-arginine signal peptide. Using both in vivo and in vitro approaches, the authors demonstrated that a sufficient amount of MPIase is required for Tat-dependent protein translocation. Furthermore, the authors successfully reconstituted the Tat transport system by combining recombinant TatA, TatB, TatC, MPIase, and FoF1-ATP synthase.

Strengths:

The reconstituted system clearly demonstrated the requirement for each component, as substrate translocation occurred only when all components were present. Based on these findings, the authors proposed a mechanistic role for MPIase in facilitating Tat-mediated membrane translocation. Previous studies have shown that MPIase is involved in Sec-dependent protein translocation and membrane protein integration, as well as YidC-dependent membrane insertion. The present study further demonstrated that MPIase also plays an essential role in the Tat translocation pathway. Overall, this work highlights the central importance of MPIase in bacterial membrane protein biogenesis and provides new insights into the molecular mechanism of Tat-dependent protein transport.

Weaknesses:

(1) To show the importance of the Tat system in bacterial cells, it would be good to describe in the introduction how many proteins are translocated via the Tat system.

(2) Figure 2B and D show that a sufficient amount of MPIase is important in SufI translocation. However, the reason why MPIase level was upregulated in the BL21 strain but not in the KS46 strain remains unexplained. The authors should address this point.

(3) In Figures 4A and B, the authors explain that MPIase first works as a receptor of TorA-GFP without recognizing the RR motif. This conclusion is based on the results of the fractionation assays, where "sup" indicates the cytoplasmic and periplasmic fractions, and "ppt" indicates the membrane fraction. In Figure 4B, under the TatABC+++, (RR), +MPIase condition, the substrate is secreted most efficiently via the Tat pathway and should therefore be recovered in the periplasm fraction (sup). However, the authors point out that efficiently processed substrate was recovered in the ppt fraction rather than the sup fraction. The authors should explain why this occurred.

Author response:

The following is the authors’ response to the original reviews.

Public Reviews:

Reviewer #1 (Public review):

Hanako and colleagues demonstrated that glycolipid MPIase is essential for the TAT system, and they successfully reconstituted the TAT system in vitro for the first time. This will facilitate the understanding of the mechanism of the TAT system.

My major points are listed below for the authors to consider:

(1) The authors successfully reconstituted the TAT system using the purified TatA/B/C, but the translocation efficiency was much lower than that of native INV. The authors partly attributed this to the reason that "MPIase recovery would be too low to detect the TAT activity" in the Discussion part. So, what would happen to the translocation efficiency if you added more MPIase to the reconstituted system? How about the abundance of MPIase from the INV and reconstituted proteoliposomes?

Fig. 3A shows that proteoliposomes reconstituted with active INV were inactive. However, TAT activity was observed when the proteoliposomes were fused with MPIase-containing liposomes. This indicates that MPIase was not sufficiently recovered during reconstitution. Typically, we added MPIase to proteoliposomes at 5% to phospholipids. However, increasing the amount of MPIase to exceed that of INV did not significantly increase the activity. The amounts of MPIase were stated in the text (P17, L3-4).

(2) Why were only TatC levels measured in Figure 2C, whereas the expression levels of TatA were not detected? Also, from my observation, the amount of TatC in the third lane is lower than that in the previous two lanes.

We also wanted to check the TatA levels, but the antibody was no longer available. Here, we confirmed that inactive INVs still contain Tat components.

(3) The authors should explain why the TatA/B/C ratios in Figure 3C (1:1:1) and Figure 3D (10:1:1) are inconsistent.

We examined various TatA/B/C ratios. We found that the ratio of 10:1:1 was better than 1:1:1. This was added to the text (P8, L11-16).

(4) ~30% of the fluorescence was recovered in the membrane fraction (Figure 4A) both in the functional TAT signal sequence (RR) and in the inactivating mutant signal sequence (KK), which suggests that MPIase acts as a relatively broad recognition factor. Given that MPIase does not discriminate between RR and KK, why do un-translocated substrates remain in the cytoplasm rather than non-specifically adhering to the membrane when MPIase is depleted in vivo?

Our results strongly suggest that MPIase acts as a signal sequence receptor. Therefore, in the absence of MPIase, the targeting of TAT precursors should be inhibited. Furthermore, because the TorA signal is less hydrophobic than the Sec signal, the level of nonspecific membrane binding would decrease.

Reviewer #2 (Public review):

Summary:

In this manuscript, the authors investigated the relationship between the Tat system and MPIase, a glycolipid that facilitates protein integration into the bacterial cell membrane. The TAT (twin-arginine translocation) system is a unique membrane transport machinery that exports fully folded proteins containing a twin-arginine signal peptide. Using both in vivo and in vitro approaches, the authors demonstrated that a sufficient amount of MPIase is required for Tat-dependent protein translocation. Furthermore, the authors successfully reconstituted the Tat transport system by combining recombinant TatA, TatB, TatC, MPIase, and FoF1-ATP synthase.

Strengths:

The reconstituted system clearly demonstrated the requirement for each component, as substrate translocation occurred only when all components were present. Based on these findings, the authors proposed a mechanistic role for MPIase in facilitating Tat-mediated membrane translocation. Previous studies have shown that MPIase is involved in Sec-dependent protein translocation and membrane protein integration, as well as YidC-dependent membrane insertion. The present study further demonstrated that MPIase also plays an essential role in the Tat translocation pathway. Overall, this work highlights the central importance of MPIase in bacterial membrane protein biogenesis and provides new insights into the molecular mechanism of Tat-dependent protein transport.

Weaknesses:

(1) To show the importance of the Tat system in bacterial cells, it would be good to describe in the introduction how many proteins are translocated via the Tat system.

The E. coli K12 cells possess 27 TAT substrates. This was added to the text (P2, L14).

(2) Figure 2B and D show that a sufficient amount of MPIase is important in SufI translocation. However, the reason why MPIase level was upregulated in the BL21 strain but not in the KS46 strain remains unexplained. The authors should address this point.

CdsA is a rate-limiting enzyme in MPIase biosynthesis. When MPIase production needs to increase, the cdsA promoters are activated. In KS46, however, cdsA is under the control of the arabinose promoter on a plasmid, so CdsA cannot be induced even when MPIase is necessary. This was added to the text (P6, L26-27).

(3) In Figures 4A and B, the authors explain that MPIase first works as a receptor of TorA-GFP without recognizing the RR motif. This conclusion is based on the results of the fractionation assays, where "sup" indicates the cytoplasmic and periplasmic fractions, and "ppt" indicates the membrane fraction. In Figure 4B, under the TatABC+++, (RR), +MPIase condition, the substrate is secreted most efficiently via the Tat pathway and should therefore be recovered in the periplasm fraction (sup). However, the authors point out that efficiently processed substrate was recovered in the ppt fraction rather than the sup fraction. The authors should explain why this occurred.

We found a mistake in the processing of the results for the sample of the TatABC+++, (RR), +MPIase condition, in Fig. 4B. Therefore, we remeasured the sample and corrected the figure. The relevant text was also modified (P9, L16-29). It was found that a large part of fluorescence was recovered in the supernatant fraction. We also uploaded the raw data for the fluorescence values.

Recommendations for the authors:

Reviewing Editor Comments:

Overall, I agree with the reviewers that the conclusions are reasonably well supported by the data. I have additional scientific/editorial concerns and recommendations:

(A) - Scientific:

(1) P.4 lines 10-15 and Fig. 1 top-left (the effect of overexpressed TatABC).

The authors state that the extent of Surf1 maturation increased upon overexpression of TatABC, but the band intensities for mature Surf1 in the TatABC-induced vs uninduced lanes are very similar.

The expression was weakened: ' The level of the mature form increased....' was changed to ' The level of the mature form slightly increased....'

(2) P.7 lines 23-24, Figure 3A. (the TAT system reconstituted in liposomes).

The authors state that a small portion of Surf1 is successfully translocated and protected by ProK. But unlike the assay results on IMVs, the size of the mature form (i.e., translocated form) appears to be the same as the untranslocated form. Some explanation seems to be needed.

In INV, the signal sequence is cleaved off by Lep, however, the cleavage is not coupled with translocation. In the reconstituted proteoliposomes, such cleavage is hardly observed since the membrane proteins were diluted by fusing with MPIase-containing liposomes. This was added to the text (P7, L26-29).

(3) Based on the model (Figure 5), the interaction between the RR motifs and MPI appears non-electrostatic as the substrate binds to the sugar moieties of the MPIase, not to the pyrophosphate part. Then what is the molecular nature of the substrate-MPI interaction? Some explanation/speculation seems to be needed.

MPIase has been identified as a factor that drives membrane protein insertion. Through the analysis of the mechanisms of insertion, we found that the glycan chain interacts directly with the transmembrane region of the membrane proteins through the numerous acetyl residues on the glycan. Moreover, we found that the positive charges of substrate membrane proteins interact with the pyrophosphate residue through the electrostatic interaction. Therefore, it is reasonable that the h region of the TAT signal binds to the MPIase glycan through the hydrophobic interactions, and the n region including the RR motif binds to the pyrophosphate residue. This was added to the text (P8, L25-27; P11, L14-17).

(B) - Editorial:

(1) P.2. lines 5-10. (Introduction)

The description of the TAT-targeting signal needs to be more clearly described for a broader readership. For example, what are "h" and "c"?

The signal sequence is composed of three regions, 'n', 'h', and 'c' from N-terminus. The n region contains positive charges including the RR motif. The next h region contains a hydrophobic stretch. The c region contains a cleavage site. This was added to the text (P2, L4-11).

(2) P.7 lines 23-24.

What is the rationale for using Pm-Fob-His as a control?

This is a control for a membrane protein unrelated to the TAT system to reveal that MPIase is specifically interacts with TatABC.

(3) P.8 lines 7-8.

CCCP is not defined.

CCCP (Carbonyl cyanide 3-chlorophenylhydrazone) is a protonophore. This was added to the text (P8, L18).

Reviewer #1 (Recommendations for the authors):

(1) The conclusions derived from Figures 1-3 heavily rely on representative immunoblotting images. Given that representative tracks can inherently introduce selection bias, how do the authors ensure the statistical robustness of these findings without quantitative bar graphs and rigorous significance analysis? To fully substantiate these interpretations, it is recommended that these immunoblotting assays be quantified across at least three independent biological replicates.

We did not quantify the results in sections where we presented qualitative discussions. However, we performed all experiments at least three times.

(2) Is it possible that the error values be added after the statistical values of translocation efficiency in Figures 2 and 3?

We added the SD values to some results in Fig. 3.

(3) "MPIase depletion of was then confirmed", an extra "of".

Corrected.

(4) The capitalization style of "TAT" in the whole text is suggested to be unified.

Corrected.

Version 2

Abstract

TAT (Twin-Arginine Translocation) is a preprotein translocation system dedicated to membrane translocation of prefolded proteins in plants and bacteria. The TAT translocon, consisting of the TatABC subunits, drives translocation using proton motive force. However, there have been no reports on the successful reconstitution of the TAT system. In this report, we show that MPIase, a glycolipid that has been known to catalyze membrane protein integration, is essential for the TAT system. Our findings in recombinant Escherichia coli demonstrate that overproducing TatABC increases MPIase levels and that depleting MPIase results in TAT precursor accumulation in the cytosol. Furthermore, co-reconstitution of MPIase with TatABC revealed the translocation activities of TAT substrates in a proton motive force-dependent manner. This is the first successful reconstitution of the TAT system and will be advantageous for understanding its mechanisms.

Introduction

The TAT (Twin-Arginine Translocation) system, found in plant chloroplasts, bacteria, and mitochondria of some protozoa and plants, is dedicated to the membrane translocation of presecretory proteins with a TAT signal sequence 1–4. This signal sequence is composed of three regions, n, h and c regions from N-terminus: The n region contains the consensus motif of “SRRxFLK”, which includes the consecutive twin arginine residues that are essential for TAT translocation 5. The central h region (hydrophobic region) of the TAT signal sequence is in average less hydrophobic than that of the Sec signal sequence 6. The c region of the TAT signal sequence often contains basic amino acids that have been shown to function as a Sec-avoidance motif 7, preventing recognition of the TAT substrates by the Sec pathway 8. This region contains a cleavage site by Lep. Unlike the presecretory proteins that are translocated through the SecYEG or Sec61 translocon, TAT substrates maintain their folded structures during translocation. In many cases, TAT substrates fold and are transported with specific cofactors 5,9. It is known that E. coli K12 possess 27 TAT substrates 9. TAT substrates are translocated through the translocation machinery formed by TatABC using the proton motive force (PMF) as the driving force 2,10–12. The TAT machinery consists of the TatA and TatC families, all of which are membrane proteins 13. In Escherichia coli, the TatA family includes TatA and TatB, as well as TatE 14–16. TatE is a functional homologue of TatA that is not essential for TAT translocation 14,17. TatA and TatB are structurally similar but functionally distinct, since both are essential for the TAT translocation 18–20. On the other hand, TatA and TatE have overlapping functions in the TAT system 14. All three proteins are integral membrane proteins containing a single transmembrane helix (TMH), followed by a cytoplasmic amphipathic helix (APH) 21,22. TatC is a polytopic membrane protein with six TMHs 23,24. TatC forms a complex with TatB at an equal molar ratio, which recognizes the signal peptides of TAT substrates 25–30. In contrast, TatA forms a homomultimer that associates with the TatBC complex with different stoichiometries depending on the size of the substrate 22,31–36. It is also reported that signal peptide alone can induce TatA multimerization 31. Binding of substrates to the TatBC complex causes TatA to assemble around the complex in a PMF-dependent manner 28,37. The TatA cluster then forms a translocon through which the TAT substrates can pass when bound to the TatBC complex 28,37–40. According to MD simulations, the short TMHs and APHs of the TatA cluster have been shown to thin the membranes, thereby destabilizing the lipid bilayer of the cell membranes and allowing the translocation of prefolded proteins 41,42. Once translocation is complete, the substrate is released from the TatABC complex, and the signal sequence is cleaved by membrane-located leader peptidase 10,43. Assembly of the TAT machinery is a transient process that occurs exclusively during substrate translocation 37,44. As mentioned above, extensive studies performed by numerous researchers have significantly advanced our understanding of the molecular mechanisms underlying of the TAT system, particularly following the identification of a Tat component in 1997 45. However, a reconstitution system for the TAT pathway has yet to be reported. Therefore, we hypothesized that a missing piece is necessary for successfully reconstituting the TAT system.

MPIase (Membrane Protein Integrase) is a glycolipid that catalyzes the integration of membrane proteins 46. Its glycan chain consists of approximately ten repeats of three N-acetylated amino sugars and is linked to diacylglycerol (DAG) by a pyrophosphate linker 46 (Figure S1). This glycan chain has a molecular chaperon-like function, i.e., the glycan interacts directly with the substrate membrane proteins prior to insertion to prevent aggregation 46–48. The glycan is flexible on the membrane surface 49. DAG in the membrane bilayer blocks the disordered, spontaneous insertion by reducing the mobility of the acyl chain within the bilayers, while MPIase restores the mobility, allowing insertion with the substrate binding 49. In combination with the pyrophosphate linker, MPIase perturbs the membrane structure, enabling membrane proteins to insert into membranes. The negatively charged pyrophosphate of MPIase also attracts the positive charges near the TMH of the substrates 47,48.

MPIase can drive the membrane insertion of simple membrane proteins with a single TMH. MPIase also acts cooperatively with proteinaceous factors responsible for membrane insertion, such as the SecYEG translocase and the YidC insertase, in both Sec-dependent and - independent membrane insertion 50,51. MPIase is co-purified with SecYEG 52 and YidC 53. , indicating that MPIase interacts with these proteins directly. MPIase also stimulates protein translocation by affecting the dimer structure of SecYEG 52,54,55. Thus, MPIase is generally involved in both membrane integration and translocation 56,57, leading us to hypothesize that MPIase is also involved in the TAT pathway.

Results

MPIase is involved in the TAT pathway in vivo

Recently, we found that CdsA and its paralogue YnbB, CDP-DAG synthases, are involved in MPIase biosynthesis and constructed the strain KS23 (ΔcdsA::cat ΔynbB) in which both the cdsA and ynbB genes were disrupted 58. The expression level of MPIase in this strain can be controlled by the addition of arabinose, since the cdsA gene is placed under the control of the araBAD promoter on a low-copy plasmid (pAra-CdsA). When arabinose is removed from the culture, MPIase becomes depleted (’ΔMPIase’ conditions) 58. We investigated the effect of MPIase depletion on the TAT pathway by using KS23/pAra-CdsA and its parent strain (EK413) (Figure 1). When the TAT substrate SufI was expressed in EK413, the signal sequence-cleaved mature form of the protein was observed, in addition to the precursor. The level of the mature form increased when TatABC was provided from the tet promoter on a plasmid (pTet-TatABC) in addition to chromosomal expression. This finding is consistent with a previous report 59. The amount of mature SufI did not change upon SufI induction, indicating that the level was saturated. A large part of untranslocated precursors was aggregated (Figure S2). In marked contrast, no mature form was observed upon MPIase depletion in KS23/pAra-CdsA, even when TatABC was expressed from pTet-TatABC (Figure 1A, top). Upon SufI induction, a smaller band marked by a star appeared in addition to the precursor; however, this band is likely a degradation product of the precursor that had not been translocated, as it is larger than the mature band. The addition of arabinose restored MPIase expression (Figure 1A, middle left) and generated the mature form of SufI (Figure 1A, top). These results strongly suggest that MPIase is involved in TAT translocation. Similar results were obtained, when using TorA, another TAT substrate: no mature form was observed upon MPIase depletion (Figure 1A, bottom left). The low level of TAT substrates under the MPIase depletion is due to the fact that MPIase-depleted cells hardly proliferate. Although TorA translocation appears to be less efficient than SufI translocation, this might be because TorA’s folding efficiency is lower in the absence of trimethylamine N-oxide (TMAO), as previously reported 60. TatABC, which had been expressed prior to arabinose removal, was still present after MPIase depletion (see Figure 2C). To confirm that the processing to the mature form represents the TAT translocation, we examined the TAT signal mutants SufI (KK) (Figure 1A, middle right) and TorA (KK) (Figure 1A, bottom right). In these mutants, the twin arginine residues in the signal sequence were replaced with twin lysine residues, resulting in an inability to translocate across the IM. We observed that these KK mutants did not convert to the mature form, even in TatABC-overproducing strains (Figure 1A, middle). These results indicate that the TAT translocation was properly studied. We investigated the involvement of MPIase in more detail using pulse-chase experiments (Figure 1B). To facilitate TAT translocation, TatABC was overexpressed from the tac promoter on a plasmid (pTac-TatABC). In the wild-type strain harboring pTac-TatABC (Figure 1B, top left), the precursor form of SufI was efficiently converted to the mature form. No mature form was observed when the signal mutant SufI (KK) was used (Figure 1B, top right), confirming that the TAT translocation was analyzed correctly. When MPIase was depleted, no mature form was observed even after a 10-min chase (Figure 1B, bottom, ‘ΔMPIase’), confirming MPIase’s involvement in the TAT system. When MPIase synthesis was induced (Figure 1B, bottom, ‘+MPIase’), the mature form appeared, but the conversion rate was much slower than under wild-type conditions. The reason for this will be discussed later. To investigate the role of MPIase in the TAT system, we performed cell fractionation into periplasmic and spheroplast fractions (Figure 1C). When the MPIase-depleted strain expressing TatABC and SufI was fractionated, the precursor was recovered in the spheroplast fraction, and no mature form was observed. This indicates that SufI was not translocated (Figure 1C, left, ‘ΔMPIase’). Conversely, mature SufI was translocated to the periplasm, when MPIase was expressed (Figure 1C, left, ‘+MPIase’). Successful fractionation was confirmed by detecting the periplasmic protein β-lactamase and the cytoplasmic protein SecB. Similar results were obtained when TorA-GFP, in which the TorA signal is connected to GFP, was examined (Figure 1C, right), confirming that TorA-GFP is not translocated upon MPIase depletion. These cells expressing TorA-GFP were also used to detect the subcellular localization directly by monitoring its fluorescence (Figure 1D). In the presence of MPIase, the envelope fraction exhibited green fluorescence, indicating that GFP is localized in the periplasm or on the inner membrane at least (left). Conversely, in the absence of MPIase, the entire cell exhibited uniform fluorescence, indicating that TorA-GFP is localized in the cytoplasm under these conditions (right). These results are consistent with those of the subcellular fractionation (see Figure 1C). Taken together, we conclude that MPIase is involved in the TAT system in vivo.

MPIase depletion causes the accumulation of TAT substrate precursors.

(A) Detection of SufI and TorA precursors upon MPIase depletion. SufI (top) and TorA (bottom) were expressed in the indicated strains. IPTG (1 mM) was added at early log phase, and cultivation was continued for 2 h. The cells were subjected to TCA precipitation, followed by SDS-PAGE/immunoblotting. The precursor and mature forms were then detected as indicated. MPIase depletion was confirmed in KS23 (middle left). Defects in the processing of SufI (middle right) and TorA (bottom right) with the KK signal were also confirmed. The symbol ’ ’ denotes degradation products of the SufI precursor. (B) Pulse-chase experiments of SufI processing after MPIase depletion. EK413/pTac-TatABC harboring pTet-SufI(RR) (top left) or pTet-SufI(KK) (top right), and KS23/pAra-CdsA/pTac-TatABC harboring pTet-SufI(RR) (+MPIase; bottom left, ΔMPIase; bottom right) were pulse-labeled for 30 s and then chased for the indicated periods. SufI(RR) and SufI(KK) were isolated using the attached His tags, followed by SDS-PAGE/autoradiography. (C) Subcellular localization of SufI and TorA-GFP after MPIase depletion. KS23/pTet-TatABC/pTac-SufI(RR) (left) and KS23/pTet-TatABC/pTac-TorA-GFP(RR) (right) were cultivated in the presence (+MPIase) or absence (ΔMPIase) of arabinose. The TAT substrates were leaky expressed from the tac promoter. The cells were then fractionated into periplasmic (peri) and spheroplast (sph) fractions. The indicated proteins were detected by immunoblotting. (D) Subcellular localization of TorA-GFP after MPIase depletion as revealed by fluorescence microscopy. The indicated strains were cultivated, TorA-GFP was induced, and then it was observed by fluorescence microscopy. The scale bars represent 10 μm. A magnified view of part of the ’+MPIase’ sample is shown on the left.

TAT substrates are translocated into INV in vitro when TatABC and MPIase are both overproduced.

(A) Translocation of SufI and TorA-GFP into TatABC-overproducing INV. SufI (top) and TorA-GFP (bottom) were synthesized in vitro in the presence of INV prepared from the indicated strains. INV was omitted in the ’-’ samples. The signal sequences on the left are wild-type (RR), while the sequences on the right are mutants (KK). Translocated materials were generated after PK digestion (+). One-fifth of the reaction mixture was analyzed as the translation standard (-). Translocation activities, expressed as a percentage of the amount of the synthesized proteins, are shown at the bottom of each autoradiogram. (B) Depletion of MPIase resulted in the failure of TAT translocation. INV were prepared from KS46/pAra-CdsA harboring pT7-TatABC, which was cultivated in the presence (+MPIase) or absence (ΔMPIase) of arabinose. After washing arabinose, IPTG was added at 1 mM to induce TatABC, and cultivation was continued for 2 h prior to INV preparation. SufI translocation was then performed as described in (A). (C) The level of TatC in INV used in (A) and (B). Each INV (10 μg) was analyzed by SDS-PAGE, followed by immunoblotting using anti-RpmJ-His antibodies. The position of TatC-His was shown. (D) MPIase is upregulated upon TatABC overproduction. The expression levels of MPIase in the strains used in (B) were determined by immunoblotting. CdsA and TatABC were induced as indicated. SecB levels were determined as a loading control. (E) MPIase is co-purified with TatABC. INV prepared from BL21/pT7-TatABC-His induced with 1 mM IPTG (+++) or not induced (-), and BL21/pT7-Pm-F0b-His induced with 1 mM IPTG, solubilized with DDM, were applied to a TALON column, then eluted. The eluates were analyzed by TLC, followed by immunodetection of MPIase. The positions of the origin/front and MPIase are shown. Purified MPIase was also analyzed at right.

Upregulation of MPIase upon TatABC overproduction is essential for in vitro TAT translocation

The TAT pathway can be reproduced in vitro using inverted membrane vesicles (INV), as depicted in Figure S3. However, it has been shown that overproduction of TatABC is necessary for translocation into INV 10,59. To reproduce these observations, we prepared INV from wild-type strains with an empty vector (Figure 2A BL21/pT7 (empty)) and from TatABC-overproducing strains (Figure 2A, BL21/pT7-TatABC), which were then subjected to the SufI translocation assay in vitro. TatABC overproduction with this vector was confirmed by immunoblotting (Figure S4). As reported 10,59, SufI translocation into INV was observed only when TatABC was overproduced (Figure 2A, top left). In the absence of INV, SufI was completely digested by PK (-). These results indicate that SufI is translocated in a TatABC-dependent manner. SufI (KK) was not translocated, even when TatABC-overproducing INV were used (Figure 2A, top right), indicating that the TAT pathway is adequately analyzed in our experimental system. We observed the same trend with another TAT substrate, TorA-GFP (Figure 2A, bottom). To further test the involvement of MPIase in the TAT pathway in this in vitro system, we used a cdsA/ynbB knockout strain with a BL21 (DE3) background (KS46) harboring pAra-CdsA. We then prepared MPIase-depleted INV with TatABC overproduction, followed by SufI translocation (Figure 2B). Since TatABC was overproduced prior to MPIase depletion, these INV contain the overproduced amount of TatABC (Figure 2C) but not MPIase (Figure 2D). We observed that SufI was not translocated into these MPIase-depleted INV, which further supports the involvement of MPIase. Consistently, the anti-MPIase antibody inhibited SufI translocation (Figure S5). However, we also found that the translocation activity with MPIase-induced INV (Figure 2B, ‘+MPIase’) did not recover to the level observed with the wild-type INV with TatABC overproduction (Figure 2B, ‘BL21/pT7-TatABC’). The reason for the incomplete recovery was that TatABC overproduction caused MPIase upregulation (Figure 2D). Significant MPIase upregulation occurred when TatABC overproduction was induced in BL21(DE3). However, MPIase expression remained unchanged with TatABC overproduction in KS46, in which cdsA was induced from the plasmid pAra-CdsA (Figure 2D). Therefore, authentic promoters for cdsA should be responsible for this upregulation, because CdsA is a rate-limiting enzyme for MPIase biosynthesis 58. Since the levels of Tat components in BL21(DE3) and KS46/pAra-CdsA grown with arabinose were similar (Figure 2C), it is reasonable to conclude that MPIase upregulation is necessary for the translocation activity in vitro of the TAT pathway into INV. This explains the low activity of MPIase-induced INV (Figure 2B, ‘KS46/pAra-CdsA +MPIase’). These results are consistent with the fact that the SufI translocation was not fully restored in the pulse-chase experiments when MPIase was induced from pAra-CdsA (see Figure 1B). MPIase upregulation was specific to TatABC overproduction; MPIase levels were barely affected by SecYEG overproduction 52 (Figure S6). These results strongly suggest that TatABC and MPIase function cooperatively. To examine whether the interaction between TatABC and MPIase is direct and physical, we detected MPIase in the TatABC-His preparation purified via metal affinity chromatography. As shown in Figure 2E, MPIase was co-purified with TatABC. Furthermore, the amount of MPIase precipitated with TatABC increased significantly with TatABC overproduction (Figure 2E). When His-tagged F0b from Propionigenium modestum (Pm-F0b-His) was expressed and used as a control, no detectable amount of MPIase was precipitated. Under these conditions, both TatABC-His and Pm-F0b-His were purified to homogeneity (Figure S7). These results indicate that TatABC and MPIase directly and specifically interact with each other. Based on these results, we concluded that the TatABC requires MPIase to function.

MPIase is the missing piece for successfully reconstituting the TAT pathway

Results obtained thus far, both in vivo and in vitro, demonstrate the essentiality of both TatABC and MPIase. Therefore, we assumed that the TAT reconstitution might be possible if TatABC and MPIase were both present and PMF was imposed. To test if our reconstitution system would work, we started with a crude system using proteoliposomes reconstituted from the solubilized membranes. First, we prepared proteoliposomes from solubilized membranes with overproduced amounts of TatABC and purified F0F1-ATPase from thermophilic bacteria (see Figure 3B, left), followed by inclusion of DAG for stable generation of PMF 61. However, these proteoliposomes were inactive with regard to SufI translocation (Figure 3A, top left, ‘+PL/DAG’).

Reconstitution of the TAT system.

(A) Membrane fusion of TatABC proteoliposomes with MPIase liposomes enabled the detection of the SufI translocation. INV, prepared from BL21/pT7-TatABC, were solubilized with OG, and then mixed with purified F0F1-ATPase. The proteoliposomes were then formed by removing the detergent. PL/DAG liposomes or PL/DAG/MPIase liposomes were fused to the reconstituted proteoliposomes. SufI translocation was performed co-translationally (left) or post-translationally (right) in the presence of these proteoliposomes. SufI(RR) (top left) or SufI(KK) (bottom left) was used as the substrate. One-ninth of the reaction mixture was analyzed as the translation standard (-) for the co-translational reactions. For the post-translational reactions, ATP was added as indicated. Translocation activities are shown at the bottom of the left panels. (B) Purification of F0F1-ATPase and the TAT components. F0F1-ATPase from a thermophilic Bacillus PS3 (left), and E. coli TatA-His (lane 1), TatB-His (lane 2), and TatC-His (lane 3) (right) were subjected to SDS-PAGE, followed by Coomassie Brilliant Blue staining. (C, D) Complete reconstitution of the TAT system using the purified components. (C) Purified TatA, TatB, and TatC (in a 1:1:1 ratio) and F0F1-ATPase were mixed with phospholipids. Then, detergent was removed to form proteoliposomes. The reconstituted proteoliposomes were then fused with PL/DAG liposomes or PL/DAG/MPIase liposomes. Translocation activity was assayed with SufI(RR) (left) or SufI(KK) (right). (D) Proteoliposomes were reconstituted with the indicated components, followed by TorA-GFP translocation. The TatA:TatB:TatC ratio was 10:1:1. CCCP (20 μM), a protonophore, was added to the sample on the left. TorA-GFP(RR) (top) or TorA-GFP(KK) (bottom) was used as the substrate. Translocation activities are shown at the bottom of each gel.

We found that fusing MPIase-containing liposomes to the proteoliposomes greatly enhanced SufI translocation (Figure 3A, top left, ‘+PL/DAG/MPIase’). This may be because MPIase was not efficiently recovered in the proteoliposome fraction due to the hydrophilic nature of MPIase’s glycan moiety 46. These results again indicate that MPIase is essential for the TAT system. Here, we could not observe the signal sequence processing. This may be due to the dilution of membrane proteins, including Lep, in proteoliposome membranes since signal cleavage and translocation are not coupled 39. Even when MPIase-containing liposomes were fused, SufI (KK) was not translocated into the proteoliposomes (Figure 3A, bottom left). We also examined the post-translational translocation using a partially purified SufI precursor (Figure 3A, top right). We observed SufI (RR) translocation into TatABC/MPIase proteoliposomes in an ATP-dependent manner (i.e., a PMF-dependent manner), but not SufI (KK) translocation. Thus, the TAT reconstitution was achieved through a combination of TatABC, MPIase and F0F1-ATPase.

Next, we purified TatA, TatB and TatC individually through an attached His tag (Figure 3B, right). We confirmed that expressing each His-tagged TatA, TatB, and TatC complemented the SDS-sensitive phenotype of the corresponding knockout strain (Figure S8). Using these purified proteins, we reconstituted proteoliposomes containing TatABC and F0F1-ATPase. Then, we fused them with MPIase/DAG liposomes. The TatABC-MPIase proteoliposomes containing F0F1-ATPase exhibited activity with respect to SufI translocation (Figure 3C, left). This activity was comparable to that of proteoliposomes reconstituted with solubilized membranes (Figure 3A). On the other hand, SufI (KK) was not translocated into the proteoliposomes (right). In these experiments, reconstituted proteoliposomes were prepared using a TatA/B/C weight ratio of 1/1/1. To obtain higher activity, we examined various TatA/B/C ratios. We found that increasing the amount of TatA gave slightly higher activity. We then prepared proteoliposomes containing TatA/B/C at a ratio of 10/1/1. Similar results with improved activity were obtained when using TorA-GFP (Figure 3D). Proteoliposomes containing both TatABC and MPIase allowed for the translocation of TorA-GFP (RR) but not TorA-GFP (KK). The translocation of TorA-GFP (RR) was severely inhibited by CCCP (carbonyl cyanide 3-chlorophenylhydrazone), a protonophore, even if the reaction mixture contains DTT that inactivates CCCP, indicating that the translocation is dependent on PMF (right). Furthermore, omitting MPIase, TatABC or F0F1-ATPase resulted in a loss of translocation activity (left). These results indicate that TatABC, MPIase and PMF are necessary and sufficient for TAT translocation.

MPIase serves as the TAT signal receptor

Although the participation of MPIase in the TAT pathway has been demonstrated both in vivo and in vitro, the mechanisms by which MPIase contributes to TAT translocation remains unclear. Our previous study showed that the glycan chain and the pyrophosphate residue of MPIase captures the substrate membrane protein on the membrane surface before membrane insertion 48,56. Although the TatBC complex has been shown to act as the TAT signal sequence receptor 26,30, we hypothesized that MPIase could also be involved in the targeting of the TAT substrate proteins to the membrane. Thus, we examined whether or not MPIase is involved in the membrane localization of TorA-GFP. Cells expressing TorA-GFP were fractionated into soluble (cytosol plus periplasm) and membrane fractions. Then, we measured the GFP fluorescence. We confirmed that the cell debris and microaggregates could be removed by low-speed centrifugation before membrane sedimentation (Figure S2). When we induced MPIase in KS23 cells by adding arabinose, ∼30% of the fluorescence was recovered in the membrane fraction (Figure 4A, ‘RR, +MPIase’), similarly to the ‘EK413’ case. Note that the MPIase level was insufficient under these conditions, resulting in very low translocation efficiency (see Figure 1B). Conversely, when MPIase was depleted, only <10% of the fluorescence was recovered in the membrane fraction (Figure 4A, ‘RR, ΔMPIase’). Similar results were obtained using TorA-GFP with a mutated signal sequence (KK) (Figure 4A). These results strongly suggest that MPIase serves as a receptor for the TAT signal, but that MPIase does not discriminate the functional signal. Similar to the ‘+MPIase’ and ‘EK413’ cases, ∼30% of the fluorescence was recovered in the membrane fraction when the tatC knockout mutant was used (Figure 4A, ‘ΔTatC’). These results suggest that MPIase recognizes the TAT signal before TatC does. Since TatABC overproduction causes MPIase overproduction (see Figure 2D), we examined the effect of TatABC/MPIase overproduction on TorA-GFP localization (Figure 4B). The amount of soluble fraction increased (∼60%), reflecting the translocation of TorA-GFP into periplasm when EK413 (cdsA+) cells, in which both TatABC and MPIase were overproduced, were used (Figure 4B, ‘RR, EK413’). In contrast, when MPIase was depleted, only 10∼20% of TorA-GFP was recovered in the membrane fraction, despite TatABC overproduction (Figure 4B, ‘RR, ΔMPIase’). These results confirm that MPIase serves as a receptor for the TAT signal. When KS23 cells with TatABC overproduction were cultivated in the presence of arabinose to induce MPIase, the amount of membrane-targeted TorA-GFP increased significantly (∼50%; Figure 4B, ‘RR, +MPIase’) compared to the absence of TatABC overproduction (∼30%; Figure 4A), indicating that TatABC stabilizes membrane-bound TAT precursors, consistent with previous reports 59. Similar results were obtained when using the KK mutant of TorA-GFP (Figure 4B, right half), indicating that MPIase cannot discriminate against the functional signal sequence. The level in the cdsA+ conditions (Figure 4B, ‘KK, EK413’) was similar to that with TorA-GFP (RR) (Figure 4B, ‘RR, EK413’). However, since this KK mutant is unable to be translocated, these results suggest that TatABC excluded the KK mutant. When GFP without the TAT signal was expressed, >90% of fluorescence was recovered in the supernatant fraction, regardless of MPIase levels and TatABC overproduction (Figure S9), indicating that the membrane localization depends solely on the TorA signal.

MPIase is a membrane receptor for TAT substrates.

(A-C) TorA-GFP is targeted to the membrane in an MPIase-dependent manner. TorA-GFP(RR) (A, B; left half) or TorA-GFP(KK) (A, B; right half) was expressed in the specified strains. The cells were then disrupted to obtain the supernatant (sup; cytosol and periplasm) and the precipitate (ppt; membranes) fractions. The fluorescence intensity of each fraction was measured. The ratio to the whole cell was determined and shown as percentages. TatABC was overexpressed in (B), but not in (A). The means and standard deviations were calculated from at least three independent experiments. *, P<0.05; **, P<0.01. In (C), the indicated strains were cultivated, TorA-GFP was induced, and then it was observed by fluorescence microscopy. The scale bars represent 10 μm. Part of each picture was magnified on each side. Scale bars represent 10 μm. (D) TorA-GFP interacts with INV expressing MPIase. Purified TorA-GFP was mixed with the indicated INV, followed by fluorescence detection (top). As a control, each INV was mixed with NBD-PE, a fluorescent phospholipid, and observed (bottom). Scale bars represent 10 μm.

The MPIase-dependent membrane targeting was also observed using fluorescent microscopy (Figure 4C). When MPIase was induced by adding arabinose, an envelope-shaped fluorescence signal was detected, regardless of whether it was wild-type or mutant (Figure 4C, left), indicating that TorA-GFP (RR) and TorA-GFP (KK) were both targeted to membranes. Conversely, under the MPIase-depleted conditions, whole-cell fluorescence was detected (Figure 4C, right), indicating that both TorA-GFP (RR) and TorA-GFP (KK) located in the cytosol.

MPIase-dependent targeting was confirmed in vitro (Figure 4D, top). When the purified TorA-GFP (Figure S10) was mixed with wild-type INV (Figure 4D, ‘EK413’), TatABC-overproducing INV (Figure 4D, ‘TatABC+++’) and TatC-deficient INV (Figure 4D, ‘ΔTatC’), in all of which contain MPIase, fluorescent dots were observed, indicating that TorA-GFP directly interacts with these INV. The number of dots increased with TatABC overproduction, which is consistent with MPIase upregulation induced by TatABC overproduction. In contrast, no such fluorescent dots were observed in the presence of MPIase-depleted INV (Figure 4D, ‘ΔMPIase’), confirming the MPIase-dependent targeting of TorA-GFP. The presence of similar amounts of membrane vesicles was visualized by adding fluorescent phospholipids, NBD-PE (Figure 4D, bottom).

Discussion

In this study, we demonstrated that glycolipid MPIase is essential for the TAT system. All of the in vivo, in vitro and reconstitution studies supported this conclusion. Therefore, the previous unsuccessful attempts at TAT reconstitution were likely due to an insufficient amount of MPIase in TatABC proteoliposomes. Even when using solubilized membranes containing TatABC and MPIase to prepare proteoliposomes, MPIase recovery would be too low to detect the TAT activity.

MPIase is a glycolipid that catalyzes the membrane insertion of proteins with a single TM 50,53 as well as those with multiple TM domains 51,55,57. Therefore, depleting MPIase should impair the membrane insertion of all the TAT components, TatA, TatB and TatC, leading to the failure of TAT translocation. However, in our experimental system, TatABC was still present after MPIase depletion (see Figure 2C), which supports the direct involvement of MPIase in the TAT translocation. Previous studies have reported that the Sec translocation slows down but does not stop completely, when MPIase is depleted 58, suggesting that the Sec machinery, including SecYEG, remains present during the initial stages of MPIase depletion, as demonstrated in this study. The results of the reconstitution study, in which TatABC was reconstituted in proteoliposome membranes, rule out the indirect involvement of MPIase in TAT translocation. In this system, the TAT translocation occurred in TatABC-, PMF-and MPIase-dependent manners.

How is then MPIase involved in the TAT system? Since MPIase has been identified as a factor that catalyzes membrane protein integration 46,50, it is possible that MPIase drives the insertion of the TAT signal sequences into membranes, thereby initiating the translocation reaction (Figure 5, (i)). Depleting MPIase caused TorA-GFP to accumulate in the cytosol. Thus, MPIase serves as a receptor for TAT precursors, and the TAT signal sequences are inserted into membranes. In this process, both hydrophobic interaction between the h-region and the acetyl residues of the MPIase glycan, and electrostatic interaction between the positive charges of the n-region and the pyrophosphate residue of MPIase would be involved, as observed in the interaction between substrate membrane proteins and MPIase 47. Then, they are transferred to TatC (ii), where the functional TAT signal is verified (iii). In addition to its roles as a membrane receptor and an insertase for the TAT signals, MPIase may facilitate structural changes in TatABC, allowing the complex to form the translocon dedicated to the substrate protein (iv). The TAT substrate is then translocated through the translocon, and the mature protein is released into the periplasm upon Lep-mediated cleavage of the signal sequence 10,43 (v). These processes require PMF, as previously reported 62. Thus, MPIase may be involved not only in the membrane targeting of TAT substrates, but also in forming the TatABC translocon for each substrate. Membrane lipids PE may be important for these processes, since PE is essential for TAT translocation 63. In the reconstitution studies, the absence of DAG in (proteo)liposomes allowed for the spontaneous insertion of membrane proteins and signal sequences 55,64. Therefore, the failure of the TAT reconstitution strongly suggests that MPIase has functions beyond serving as a membrane receptor and an insertase for TAT translocation. Given that MPIase induces dynamic structural changes in the SecYEG translocon 52,54, it is likely that MPIase also modifies the structure of TatABC to form a tailor-made translocon. This possibility is consistent with the observations showing that TatABC overproduction induces MPIase upregulation and that MPIase co-precipitates with TatABC.

Working model of mechanisms underlying MPIase-dependent TAT translocation.

MPIase recognizes the signal sequence of the TAT substrates and insert them into the membrane. The TatBC complex on the membrane then determines whether the signal sequence of the membrane-inserted TAT substrate is valid. The TatABC translocon then forms a pore the same size as the substrate. This allows the substrate protein to be translocated across the membrane. Finally, Lep cleaves the signal sequence, and the mature protein is released into the periplasm. See the text for details.

It is well known that the lipid environment of the membranes significantly affects the efficiency of TAT translocation. First, acidic phospholipids are important for TAT translocation because TAT signals interact with the acidic phospholipids 65. MPIase has a pyrophosphate residue and would therefore contribute to this interaction in a similar way. At the same time, the hydrophobic domain of the TAT signals is solubilized by the glycan chain, as occurs with membrane proteins 47. Conversely, a decrease in PE (i.e., an increase in acidic phospholipids, PG and CL) decreases efficiency 66,67, presumably due to an overly strong interactions between the TAT signals and the acidic phospholipids. Alternatively, PE may be important for TAT translocation, since it stabilizes several membrane proteins 68. Thus, PE may be involved in the expression of the TatABC function. In the thylakoid system, the reduced unsaturation in membrane lipids results in a 50% decrease in the efficiency of TAT translocation 69, suggesting that membrane fluidity is also important for TAT translocation. We have demonstrated that MPIase increases the membrane fluidity through the flexible movement of its long sugar chain 49. This property would benefit TAT translocation.

MPIase is upregulated in the cold. During this process, both cdsA and ynbB are induced at the transcriptional level. In the case of cdsA induction, two cold-inducible promoters located upstream of the cdsA gene are responsible 70. MucA, a small peptide that overlaps with cdsA, enhances MPIase synthesis by CdsA 71. When MPIase is upregulated by TatABC overproduction (see Figure 2D), cdsA and/or ynbB are also induced, as evidenced by the diminished increase in the MPIase level in cdsA/ynbB knockouts even when CdsA expression is complemented from the araBAD promoter on a plasmid. In this case, the cdsA promoters other than identified cold-inducible ones would be responsible, together with MucA. Alternatively, other mechanisms could operate to upregulate MPIase.

We found that the TAT translocation could be reconstituted with TatABC and MPIase alone when PMF was imposed, indicating that these factors are sufficient for TAT translocation. Although the activity was comparable to that of INV and proteoliposomes with the crude extract, it was still low. Therefore, these results suggest the presence of factor(s) involved in TAT translocation that stimulate the reaction, as the soluble factors that solubilize the TAT substrates and target them to membranes remain unidentified.

The TAT system is widely used as a carrier system for secreting foreign proteins into the medium 72–74. Now that we know that MPIase is involved in the TAT system, it will be easier to engineer the TAT system by modulating the MPIase activity to increase the level of secreted proteins.

The TAT system was first identified in the thylakoid membranes of chloroplasts as a ΔpH-dependent pathway 2,75. The components of the TAT system in chloroplasts are highly homologous to TatABC in E. coli 26,37, and the translocation reaction is also driven by PMF 2. Thus, the mechanisms underlying the TAT systems in E. coli and chloroplasts likely have similar underlying mechanisms, strongly suggesting the presence of an MPIase homologue in chloroplasts. This similarity could facilitate the identification of an MPIase homologue in chloroplasts.

Methods

Materials and methods

The strains, plasmids, and oligonucleotides used in this study are listed in Tables S1, S2, and S3, respectively. INV were prepared as described 59. The plasmid pCspB-TorAss-GFP encoding TorA-GFP was obtained from Dr. Y. Kikuchi (Ajinomoto Co., Inc.). Bio Beads SM-2 were purchased from BioRad. [35S] EXPRE35S35S Protein Labeling Mix, a mixture of [35S] methionine and [35S] cysteine (∼37 TBq/mmol), was obtained from Revvity. Proteinase K (PK) was purchased from Roche Diagnostics. The cell-free protein synthesis kit “E. coli S30 Extract System for Circular DNA” was purchased from Promega Corporation. The E. coli polar lipid extract (phospholipids), 18:1 NBD-PE, and dioleoylglycerol were obtained from Avanti Polar Lipids, Inc. n-Octyl-β-D-glucopyranoside (OG) and n-dodecyl-β-D-maltopyranoside (DDM) were obtained from Dojindo Laboratories. IPTG was from Wako Pure Chemicals Industries, Ltd. TLC plates were from Merck. MPIase was purified from MC4100, as described 46,50. TorA-GFP, TatA, TatB, TatC, and the TatABC complex were purified from DH5α/pTac-TorA-GFP-His, DH5α/pTac-TatA-His, DH5α/pTac-TatB-His, DH5α/pTac-TatC-His, BL21/pT7-Pm-F0b-His, and BL21/pT7-TatABC-His, respectively. These proteins were induced with 1 mM IPTG, followed by INV preparation, solubilization and purification through the 6 x His-tag attached to C-termini on Talon columns. The conditions for INV solubilization and Talon column chromatography was carried out as described 57. Purification of F0F1-ATPase from a thermophilic bacterium PS3 was performed as described 76.

Depletion of MPIase

Strains KS23/pAra-CdsA and KS46/pAra-CdsA were cultivated overnight at 37°C in LB medium containing 0.002% arabinose. After the cells were washed five times with fresh LB medium, they were inoculated into LB medium supplemented or not with 0.2% arabinose at a dilution of 1/1000 and then shaken at 37°C. To express the TAT substrates, IPTG (1 mM) was added after the cell culture was grown to OD600 nm ∼0.5 and cultivated for another 1 h. An aliquot (500 µL) of each sample was treated with TCA to a final concentration of 5%. The TCA precipitate was recovered by centrifugation (10,000 x g, 4°C, 5 min), followed by washing with acetone. MPIase depletion was then confirmed by SDS-PAGE/immunoblotting, as described 58.

Pulse-chase experiments

Cell cultures grown in LB medium to OD600 nm ∼0.5 were harvested and resuspended in M9 minimal medium containing 18 amino acids (0.1 mM each) except for methionine and cysteine, and 0.1% yeast extract. After incubating for an additional 1 h at 37°C, the substrate proteins were labeled with [35S] methionine/cysteine (∼10 MBq/mL) for 30 s. Labeling was terminated by adding non-radioactive methionine and cysteine to a final concentration of 12 mM. At the specified chase time, an aliquot (500 µL) of the culture was withdrawn, and TCA was added to a final concentration of 10%. The TCA precipitates were washed successively with acetone and diethyl ether, and then suspended in 50 µL of 50 mM Tris-HCl (pH 7.5), 1% SDS, 1 mM EDTA (pH 8.0), and boiled for 3 min. The samples were then diluted with 1 mL of 50 mM HEPES-KOH (pH 7.5), 150 mM NaCl, 1% Triton X-100, 1 mM EDTA (pH 8.0), 1 mM phenylmethylsulfonyl fluoride, and were centrifuged at 16,000 x g for 5 min. Then, 10 µL of TALON resin was added to the supernatant, and the mixture was stirred at 4°C for 1 h. After centrifugation (3,000 x g, 4°C, 30 s), the TALON resin was washed successively with 500 µL of 50 mM Tris-HCl (pH 7.5), 150 mM NaCl, 1% Triton X-100. The washed resin was suspended in 20 µL of the sample buffer for SDS-PAGE. Radioactive bands, analyzed by SDS-PAGE, were detected by autoradiography using a phosphorimager (GE Healthcare), and quantified using ImageQuant (GE Healthcare).

Preparation and fractionation of spheroplasts

After the cells were grown in LB medium to OD600 nm ∼0.5, 1 mM IPTG was added to induce the TAT substrates. The cells were harvested at 2,400 x g for 5 min and washed with 5 mL of 0.75 M sucrose, 10 mM Tris-HCl (pH 7.5). They were then suspended in 1 mL of 0.75 M sucrose, 10 mM Tris-HCl (pH 7.5) and 20 µg/mL lysozyme solution and incubated on ice for 5 min. They were converted into spheroplasts by adding 1.5 mM EDTA (pH 8.0) at equal intervals over 10 min, bringing the total volume to 2 mL. An aliquot of the spheroplast suspension (1 mL) was centrifuged (15,000 x g, 4°C, 1 min) to sediment spheroplasts. The spheroplasts were suspended in 500 µL water and treated with 10% TCA, followed by incubation on ice for 10 min to precipitate proteins. The periplasmic fraction, obtained by sedimentation of the spheroplasts, was treated similarly. The proteins were then recovered by centrifugation (10,000 x g, 4°C, 5 min), washed with acetone, and subjected to SDS-PAGE/immunoblotting, as previously described 77.

Fluorescence microscopy

E. coli strains were grown in LB medium to OD600 nm ∼0.5, followed by the addition of 1 mM IPTG to induce TorA-GFP. The cells were fixed on glass slides coated with 1% agarose, as described 39. The cells were observed using a fluorescence microscope (Olympus BX53, combined with a U-FGFP filter). Images were captured with a 3CCD digital camera (Olympus DP74). The digital images were then processed and analyzed using the cell image analysis software, cellSens Standard (Olympus).

Assaying of the protein translocation in vitro

The TAT substrates were synthesized in vitro using the E. coli T7 S30 extract system. For the co-translational translocation assay, the followings were added to a 20 µL reaction mixture: [35S] methionine and cysteine (∼10 MBq/mL), plasmids encoding the substrate protein under the T7 promoter, and INV or proteoliposomes. The synthesis/translocation of the RI-labeled substrates were performed at 37°C for 1 h. When INV was used, it was added to the reaction mixture 5 min after the reaction started. For the post-translational translocation assay, the RI-labeled substrates (3 µL), synthesized in the absence of INV or proteoliposomes at 30°C for 30 min, were added to a translocation mixture containing 50 mM HEPES-KOH (pH 7.5), INV or proteoliposomes, 1 mM MgSO4, 5 mM creatine phosphate, 10 µg/mL creatine kinase, and 1 mM ATP. After incubating the mixture at 37°C for 45 min, it was divided into two portions. One portion was treated with 5% TCA to precipitate the substrate proteins. The other portion was mixed with an equal volume of PK (1 mg/mL) and incubated at 25°C for 20 min. After adding 5% TCA, the samples were incubated at 56°C for 5 min to inactivate PK. Then, the proteins were recovered by centrifugation (10,000 x g, 4°C, 5 min). Radioactive materials were detected and quantified as described above.

TALON affinity chromatography

INV (4 mg protein) was solubilized using 1% DDM, 20% glycerol, 50 mM HEPES-KOH (pH 7.5). The solubilized membranes were recovered by centrifugation (170,000 x g, 4°C, 30 min). Samples were applied to a TALON column (0.5 mL). After washing the column with 20 mL of 0.02% DDM, 10% glycerol, and 50 mM HEPES-KOH (pH 7.5), the column-bound materials were eluted with the same buffer containing 150 mM imidazole. The eluted fraction was analyzed by TLC using a solvent system C (chloroform/ethanol/water: 3/7/4) for development. Then, MPIase was detected using anti-MPIase antibody, as described 53.

Reconstitution of proteoliposomes

Phospholipids (13 mg) were resuspended in buffer A (50 mM HEPES-KOH (pH 7.5), 1 mM dithiothreitol) at 20 mg/mL, followed by bath sonication to form liposomes. The liposomes were solubilized in 5% [w/v] OG and incubated on ice for 15 min. Then, 200 mg of Bio Beads SM-2 were added, and the mixture was gently stirred at 4°C for 30 min. Purified TatA (100 µg or 1000 µg), TatB (100 µg), TatC (100 µg) and F0F1-ATPase (1 µM) were added to the solution. The solution was gently mixed at 4°C for at least 3 h. The Bio Beads SM-2 were then removed, and the proteoliposomes were recovered by centrifugation (170,000 x g, 4°C, 1 h). DAG- and DAG/MPIase-containing liposomes were prepared, as described 77. TatABC/F0F1-ATPase proteoliposomes and PL/DAG liposomes or PL/DAG/MPIase liposomes were typically mixed at a ratio of 4:1, followed by three cycles of freezing, thawing, and sonicating to allow liposome fusion, as described 54. As a result of liposome fusion, proteoliposomes contain DAG and MPIase at 5% each to phospholipids. The mixture was then passed through a mini-extruder (Avanti Polar Lipids, Inc.) with a 0.4 µm pore size filter to obtain uniformly sized liposomes, as described 64.

Subcellular fractionation

After the cells were grown in LB medium to OD600 nm ∼0.5, 1 mM IPTG was added to induce the TAT substrates. The harvested cells were suspended in 50 mM HEPES-KOH (pH 7.5), 10% glycerol, 1 mM dithiothreitol, 1 mM phenylmethylsulfonyl fluoride, protease inhibitor cocktail (Roche) and then disrupted using a French press. Then, the cells were incubated with 2 mM MgSO4, DNase I (1 mg/mL) on ice for 15 min, followed by the addition of 2 mM EDTA (pH 8.0) and incubation on ice for an additional 15 min. The disrupted and undisrupted cells were separated by centrifugation (15,000 x g, 4°C, 10 min) and then they were fractionated into soluble (cytosol and periplasm) and insoluble (membrane) fractions by centrifugation (150,000 x g, 4°C, 1 h). TorA-GFP was detected by monitoring GFP fluorescence. Samples were added to 2 mL of 50 mM HEPES-KOH (pH 7.5), and the fluorescence intensity was measured at an excitation of 395 nm and an emission of 509 nm using an FP-8200 fluorescence spectrometer (JASCO).

Statistical analysis

Statistical significance was determined using a Student’s t-test. The P values of less than 0.05 were considered significant.

Other methods

SDS-PAGE was performed using 12.5% acrylamide/0.27% N,N’-bis (methyleneacrylamide) 78 for TatA, TatB, TatC, SecB, β-lactamase, and MPIase. 15.0% acrylamide/0.33% N,N’-bis (methyleneacrylamide) 79 was used to analyze SufI and TorA-GFP, while 8% acrylamide/0.17% N,N’-bis (methyleneacrylamide) was used for TorA. Bands on the immunoblots were quantified using a CS analyzer (ATTO). Protein quantification after INV preparation was performed using bovine serum albumin as a standard, as described 80. Detection of MPIase and other proteins by immunoblotting was performed as described 58,81.